### 2011-12-12 paper.tex
edits + latex fixes

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

% preprint + pagenumbering for page numbers
\documentclass[preprint]{style}
\pagenumbering{arabic}

\usepackage{paralist}
\usepackage{graphicx}

% This gets rid of the block of whitespace on the first page
\makeatletter
\let\@copyrightspace\relax
\makeatother

\begin{document}

\title{Survey of Cache Oblivious B-trees}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Ahmad\\
\affaddr{MIT CSAIL}
\email{saahmad@mit.edu}
% 2nd. author
\alignauthor
Leilani Battle\\
\affaddr{MIT CSAIL}
\email{leibatt@mit.edu}
% 3rd. author
\alignauthor
Stephen Tu\\
\affaddr{MIT CSAIL}
\email{stephent@mit.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\date{7 December 2011}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\newcommand{\lhyperceil}{\lceil\lceil}
\newcommand{\rhyperceil}{\rceil\rceil}
\newcommand{\lhyperfloor}{\lfloor\lfloor}
\newcommand{\rhyperfloor}{\rfloor\rfloor}
\newcommand{\llangle}{\langle\langle}
\newcommand{\rrangle}{\rangle\rangle}

\newcommand{\Search}{\textsc{Search(k)}}
\newcommand{\Insert}{\textsc{Insert(k, v)}}
\newcommand{\Insertkonly}{\textsc{Insert(k)}}
\newcommand{\Scan}{\textsc{Scan(a, b)}}
\newcommand{\Delete}{\textsc{Delete(k)}}
\newcommand{\Succ}{\textsc{Succ(k)}}
\newcommand{\Pred}{\textsc{Pred(k)}}
\newcommand{\Member}{\textsc{Member(k)}}
\newcommand{\Range}{\textsc{Range-Query(k$_{i}$,k$_{j}$)}}

\maketitle
\begin{abstract}

This paper presents a survey of cache-oblivious B-trees (CO B-trees). The CO
B-tree was first introduced by Bender et al \cite{BenderDemainColton}. It is
an external memory data structure; it optimizes the number of memory transfers
performed rather than the number of CPU instructions that executed. Unlike
other external memory data structures such as the widely used B-tree, the CO
B-tree does not have access to the parameter's of the machine's memory
hierarchy, including the block size/capacity at each level, or the relative
speeds. Hence, the CO B-tree is \textit{cache-oblivious}.  Despite this, the CO
B-tree is able to achieve the same provably optimal search bound as the B-tree
and near optimal insertion and deletion bounds. In this paper we first describe
the original CO B-tree and explain how it is able to achieve its performance.
We then present and analyze three followup CO B-tree data structures that
improve upon weaknesses of the original: the \textit{cache-oblivious string
B-tree}, which introduces support for variable length keys, the
\textit{cache-oblivious streaming B-tree}, which improves the insertion bound
at the expense of the search bound, and lastly, the \textit{cache-oblivious
concurrent B-tree}, which introduces support for concurrency.

\end{abstract}

\section{Introduction}

\subsection{External Memory Algorithms}

Most algorithms assume an idealized model of computation that consists of
memory as a flat byte-addressable array which can be read and modified in some
fixed time. However, in practice memory is not a flat array, but rather is
organized as a hierarchy consisting of fast on-chip caches (L1, L2, etc.), to
slower DRAM, to even slower disks. When a program requests data, the data is
located in the memory hierarchy and typically carried up the chain to the
fastest memory layer (closest to the CPU), usually a cache. Unfortunately, this
cache can only have a limited capacity, and once it is full any new data
requests necessitate some existing cached data to be evicted. Transferring data
from lower to higher levels in the memory hierarchy often takes an
exponentially long time; access times between the different layers can differ
by many orders of magnitude. Therefore, the time to transfer data often
dominates the overall runtime performance of an algorithm.

External memory algorithms are wary of this fact and seek to optimize the
number of memory transfers performed rather than the number of CPU instructions
executed. They often try to exploit the fact the memory hierarchies transfer
data in blocks rather than individual machine word units. Thus, external memory
algorithms are designed to not only limit the number of memory transfers, but
also preserve the locality of related data so that when a costly memory access
is performed, its cost is amortized.

\subsection{B-trees}

Since external memory algorithms are primarily concerned with accessing data,
it is no surprise that many external memory problems can be reduced to finding
a suitable data structure that efficiently implement certain operations. A well
known external memory data structure is the B-tree \cite{BayerM72}. The B-tree
works by making each one of its nodes the same size as a memory block in a
particular level of the hierarchy. Furthermore, the fanout in the tree (the
number of children of a node) is proportional to the block size as well.  The
B-tree incurs $O(\log_{B}{N})$\footnote{Throughout this paper, $\log$ with the
base omitted is base-$2$} memory transfers for \Search{}, \Insert{}, and
\Delete{} and is widely used in databases and file-systems.

\subsection{DAM vs CO}

The B-tree has two theoretical shortcomings. First, the B-tree is analyzed in a
two-level memory model called the Disk Access Machine (DAM) \cite{Aggarwal}.
However, in practice, architectures typically have memory hierarchies that are
larger than two. The argument for the DAM model is that in practice, only one
level in the memory hierarchy tends to dominate the performance, and thus a
two-level memory model is sufficient.  However, this means that B-trees can
begin to incur an increasing number of memory transfers if the dominating level
in the hierarchy changes.

Additionally, the organization of a B-tree requires knowing the block size of
the memory layer in advance. B-trees are implemented with parameterized code
that needs to be changed for every level in the memory hierarchy. This tuning
can be very difficult and error prone. Additionally, tuning this parameter may
even be impossible in certain heterogeneous computing environments.

The shortcomings of the DAM model gave rise to the cache-oblivious (CO) model
of computation \cite{Frigo99}. In the CO model an algorithm has no information
about the memory hierarchy. It does not know anything about the speed or the
block size of any level. Conceptually, the CO model allows algorithm designers
to reason about the simple two-level model like DAM but also prove results for
any arbitrary multi-level memory hierarchy. This is because algorithms that
perform well in a CO model should, in theory, perform well regardless of
changes in block size, cache capacity, and cache speed. Consequently, CO
algorithm should gracefully scale to multi-level hierarchies as well.

\subsection{CO B-trees}

In \cite{BenderDemainColton}, Bender et al. introduce the cache-oblivious
B-tree (CO B-tree). The CO B-tree is similar to the normal B-tree except that
it does not require explicit parameters about the memory hierarchy.

The CO B-tree is able to achieve the optimal search bound of
$\Theta(\log_{B}{N})$ memory transfers, the optimal scan bound of $\Theta(N/B)$
memory transfers and near optimal insertion and deletion bound of 
$\Theta(\log_{B}{N} + \frac{\log^2{N}}{B})$. 
The CO B-tree was a seminal data structure that gave rise to many others.

This paper provides a survey of the different types of CO B-trees and discusses
the various trade offs that they make. We start by analyzing the original CO
B-tree and explain how it works. This includes a detailed description of the
packed memory array (PMA) and strongly weight-balance binary search trees
(SWBST) that are integral to the CO B-tree's memory performance. Following this
discussion we provide an analysis of the CO string B-tree (a CO B-tree with
variable length keys), the CO streaming B-tree (a CO B-tree that improves the
cost of insertion at the expense of search), and the CO concurrent B-tree (a CO
B-tree that allows multiple processes to efficiently access data concurrently).

\section{Background}

\subsection{Original CO B-trees}
\label{sec:original}

The CO B-tree optimizes the number of memory transfers by structuring its
elements as a strongly weight-balanced search tree (SWBST). The SWBST is then
laid out into a packed memory array (PMA) using a van Emde Boas (vEB) layout.
The SWBST and PMA play crucial roles in the other cache-oblivious data
structures that were developed after the CO B-tree and thus are described at
length in the next two sections.

\subsection{Structure - SWBST}
\label{sec:structure}

\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/veb.pdf}
\end{center}

\caption{An illustration of the van Emde Boas (vEB) layout}
\label{fig:veb}
\end{figure}

The ultimate goal of a CO B-tree is to reduce the number of memory transfers
while supporting \Search{}, \Insert{}, and \Delete{} operations. Thus, it is
not surprising that the CO B-tree maintains its data in a balanced search tree.
The search tree is laid out in memory according to a van Emde Boas (vEB) layout
\cite{veb1,veb2}. The idea of a vEB layout is fairly straight forward. Let $h$
denote the height of a tree. The vEB layout splits the tree approximately at
height $h/2$. This splits the original tree into a \textit{top} subtree and
many \textit{bottom} subtrees.  Let $A$ denote the top subtree and
$B_1,...,B_k$ denote the bottom subtrees.  The vEB layout is defined
recursively using the order: $A,B_1,...,B_k$. This layout is also shown in
Figure~\ref{fig:veb}.

A nice consequence of using a vEB layout is that it can be shown to use
$O(\log_{B}{N})$ memory transfers for searches \cite{veb1, veb2}. Thus, the CO
B-tree has already achieved the desired bound for search queries. It still
remains to be shown that the CO B-tree achieves the desired number of memory
transfers for dynamic inserts and deletes.

To support dynamic inserts/deletes, a \textit{strongly weight-balanced search
tree} (SWBST) is used \cite{swbst}.  The weight of a node $u$ is equal to the
total number of $u$'s descendants plus 1. It can also be thought of as the sum
of the weights of $u$'s children plus 1. Formally, $u$'s weight is $w(u) = 1 +
\sum_{v \in children} w(v)$.

Most weight-balanced search trees require that the heights of a node's left and
right subtrees differ by at most a constant factor. The SWBST enforces the
stronger property that a node at height $h$ in the tree (leaves have a height
of 1) has $\Theta(c^h)$ descendants where $c$ is some constant ($c > 4$). This
property becomes important later on when laying the tree out in memory (see
Section~\ref{sec:layout}).

The definition of a SWBST given in \cite{swbst} provides this bound for the
weight of a node $u$:
\begin{equation}
  \frac{c^{h-1}}{2} \leq w(u) \leq 2c^{h-1}
\end{equation}

Let us now consider insertions. To insert an element, search
down the tree and find the natural place to insert the element - just like in
any tree structure. After inserting the element into some node $w$, 
check to see if any of $w$'s ancestors have become unbalanced by having a
weight greater than $2c^{h-1}$. If such an ancestor exists, it follows that
every node along the path between that ancestor and $w$ will be unbalanced as
well. This requires splitting nodes to restore balance, starting from $w$
and moving up the tree.

When a node is split, we need to determine which children belong to which of
the two nodes. Dividing the children evenly is insufficient because it may
cause another violation. Instead, find the longest sequence of elements such
that the weight of the subsequence is $\lceil w(u)/2 \rceil$.  By doing this,
we maintain the bound that $\frac{c^{h-1}}{2} \leq w(u) \leq 2c^{h-1}$. Cascade
these changes up the tree until all of the ancestors are balanced.

Deletions are supported similarly, except we check if a node has a weight $w(u)
< \frac{c^{h-1}}{2}$. In this case, we merge two neighbors together and proceed
up the tree. There is one edge case that needs to be handled. The merging
procedure may produce a new node that has a weight larger than the bounds
allow. If this happens, immediately split the node again, recursively. This can
be thought of as stealing nodes from a neighbor.

We have now sketched that the SWBST has the desired property that the number of
descendants of a node is $\Theta(c^h)$. We now describe the tree's memory layout.

\subsection{Layout - PMA}
\label{sec:layout}

Now that the elements structured in a SWBST, we are still left with the problem
of mapping this tree to memory and proving that it achieves the desired number
of memory transfers per operation. The packed memory array (PMA) was developed
to solve this problem \cite{BenderDemainColton}. There are three main costs
that the PMA is concerned with: \textit{inserts}, \textit{splits/merges}, and
\textit{pointer updates}.

First, the PMA needs to maintain open space in its memory for newly inserted
items. The PMA is able to handle an insertion into the SWBST using only $O(1 +
(\log^2{N}) / B)$ memory transfers (this is the cost of updating the memory
layout of the SWBST in the PMA, not the cost of updating the data structure
itself). Doing so may seem contradictory: maintaining the locality of elements
while keeping space available for new inserts. The PMA uses a technique
described in \cite{packedmemoryarray}. At a high level, the PMA maintains
windows into the array. When a window becomes too full or too empty, the
elements are spread out within the context of a larger window. Two parameters
are required: the window size and the threshold for triggering a spread out.
However, these parameters are not related to the block size of the memory level
and thus do not violate the CO model.

Second, the PMA needs to handle splits and merges caused by inserting into the
SWBST discussed in the previous section. The PMA is able to accomplish this
using $O(1 + (\log{N}) /B)$ memory transfers. The proof is based on the fact
that each node has $\Theta(c^h)$ descendants (recall, this is the key property
that SWBST provided). Thus, merging or splitting a node will cause
$\Theta(c^h)$ nodes to move, requiring $O(1 + c^h/B)$ memory transfers. It
follows from the definition of SWBST in \cite{swbst} that when a node is
rebalanced, it will require $\Theta(c^h)$ insertions or deletions until it is
rebalanced again. Thus, the amortized cost of each insertion and/or deletion is
$O(1/B)$. When a node is split, it may traverse all the way up the tree, and
thus the split/merge bound is $O((\log N) / B)$ memory transfers.

Third, the PMA needs to handle updating the parent and child pointers that may
become invalidated because of how the nodes moved around. If the pointers are
all packed together in memory (e.g. within $B$ distance of each other) then
only $O(1)$ memory transfers are required. However, if they are further apart
then a problem arises. To deal with this, buffer nodes are introduced.  Buffer
nodes ``stretch'' the vEB layout so that nodes do not move far away from one
another. By using buffers, the PMA is able to achieve a pointer update cost
using $O(1+ \frac{\log{B}}{\sqrt{B}}\log^2{N})$ memory transfers.

Of these three terms, the dominating factor is maintaining room for newly
inserted elements - $O((\log^2{N})/B)$. Therefore, overall the CO B-tree
achieves insertions and deletions in $O(\log_B{N} + (\log^2{N})/B)$ memory
transfers. It still achieves the optimal $O(\log_B{N})$ memory transfers for
searches, matching the performance of static cache-aware B-trees.

\section{Variable Length Strings}
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/COSB-treegeneralstructurediagram.pdf}
\end{center}

\caption{An illustration of the dynamic COSB-tree structure}
\label{fig:string_structure}
\end{figure}
\subsection{Motivation}

In order for B-trees to remain versatile, they must be able to accommodate
various forms of data. However, traditional B-trees and string B-trees have
suboptimal performance when variable length strings are used as keys. In
addition, string B-trees do not support compression, and commonly used front
compression is handled sub-optimally. Regular B-trees and string B-trees are
not cache oblivious, and thus incur poorer performance as described
previously. Range queries have poor performance in B-trees. In
\cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache
Oblivious String B-tree (COSB-tree), which they argue works efficiently in
these problem areas for variable length string keys. They also provide a
modified front-compression scheme, which gives their approach the advantages
of using less space and having improved locality within the compressed data.

Two experiments are performed, in which a regular B-tree implementation tuned
to various block sizes is compared to a static COSB-tree implementation. The
experiment involved measuring the amount of time required to perform: 440,000
and 450,000 inserts of random values, a range query of all values, and 1000
random searches. The results of the experiment show that COSB-trees
outperformed B-trees most of the time for random searches, range queries and
random inserts. However, some point between 440,000 and 450,000 inserts the
COSB-tree had to rebalance the entire tree, causing it to fall behind in this
measurement. However, Bender et al. argue that the experiment was biased
towards the B-tree implementation, since the trees are "young", and thus still
well organized. The COSB-tree implementation is also compared to Berkeley DB
**cite berkeley db**, and achieves performance comparable to Berkeley DB with
tuned parameters. The results seem to show that there is a window in which
large and small-block size B-trees perform poorly, and that COSB-trees may be
a reasonable alternative to improve performance.

\subsection{Description}

\subsubsection{General COSB-tree Structure}

The static COSB-tree consists of 2 layers: a centroid tree at the top, and an
array of the keys stored in lexicographic order on the bottom. The leaves of
the centroid tree contain pointers to keys in the array. The centroid tree is
structured such that a search through the centroid tree identifies the local
area of the desired key in the array. Then a sequential scan is performed
across the array to find the key.

The Dynamic COSB-tree consists of 3 major components: a centroid tree at the
top; a packed memory array (PMA) structure of hash data for the keys known as
the \emph{hashdata} layer; and on the bottom a PMA structure for the original
key values, compressed using an augmented version of forward compression known
as the \emph{keydata} layer.

The centroid tree on a high level behaves similarly for static and dynamic
COSB-trees. The leaves of the centroid tree point into the hashdata layer,
where a sequential scan is performed to find the correct key by hash value.
The hashdata layer points directly into the keydata layer, where the
appropriate key is decoded, compared and returned.

In the following subsections, centroid trees are explained, and a summary of
how to modify them to preserve balance and structure in the dynamic case is
provided. Two iterations on forward compression are also summarized to show
how to incrementally add support for variable length strings as keys for
dynamic COSB-trees.

\subsubsection{The Dictionary Matching Problem and Centroid Trees} %todo:cite dictionary matching problem--sources in cosb-tree paper
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidtreeexample.pdf}
\end{center}

\caption{A centroid tree example}
\label{fig:centroid}
\end{figure}

In this approach to solving the dictionary matching problem *cite dictionary
matching problem paper*, a dictionary $D$ of $N$ keys is preprocessed using
hashing and divide-and-conquer via the construction of a centroid tree over
the keys. This technique forms the basis for the Cache Oblivious String B-tree
(COSB-tree) design. Consider the compacted trie $T$ of $D$. Bender et al.
observe that there exists a centroid vertex $\rho$ in $T$ such that $\rho$ has
between N/3 and 2N/3 descendants, inclusive. When searching for a key $k$ in
the compacted trie, you traverse the tree as follows: for each node $y$ that
you visit, if $y$ is a prefix of $x$, or $y$ matches $x$, continue on into the
trie rooted at $y$ or \emph{down trie} of $y$. Otherwise traverse the trie
resulting from excluding $y$ and its subtree. The \emph{centroid tree} is
obtained by rooting the tree at \emph{centroid vertex} $\rho$, and making its
children recursively defined centroid trees of its up and down tries.

At each step through the centroid tree, a consistent fraction of the nodes are
eliminated from the search, resulting in an $O(logN)$ time bound when ignoring
the time required to operate on keys. To make key comparisons faster, the keys
are hashed and hash values are compared in place of the original key values.
To avoid mismatches due to overlap in hash values, the original key values are
compared after finding a match in hash values. The resulting time bound for
\Search{} (interchangeable with \Member{}) is $O(||k||+\log{}N)$. If each node
in the centroid tree knows its successor and predecessor descendants, \Pred{}
and \Succ{} can be achieved in $O(||k||+||k'||+logN)$, where $k'$ is the
successor or predecessor of $k$. Note that the additive $logN$ incurred by the
use of centroid trees can be avoided when using main memory. Centroid trees
are used to capitalize on locality for the cache oblivious string B-tree
design, described later in this section.

The centroid tree is built using $O(N/\log{}N)$ keys, and the centroid tree is
layed out in memory using a modified van Emde Boas (vEB) layout for
weight-balanced binary trees. The \emph{hyperhyperfloors} of nodes are used as
a guide for cutting the tree into equal-size subtrees for the vEB layout. The
\emph{hyperhyperfloor} of x is denoted as $\lhyperfloor{}x\rhyperfloor{}$, and
refers to rounding x down to the nearest power of a power of 2. This ensures
trees of equal size on all \emph{levels of detail}, as the centroid tree is
traversed.

The following approach is used to rebalance the centroid tree: Put all the
elements in DFS or Euler tour order for the trie. Given this order, scan the
trie to identify the centroid vertex, partition the trie into its respective
up and down tries, and recurse on the up and down tries. The Euler tour order
can be achieved via $O(\log{}N)$ scans, since the trie is only $O(\log{}N)$
deep. Thus updates have amortized cost of $O(1+\log^{2}N/B)$ memory transfers.
Note that this doesn't include updating successor or predecessor pointers,
which is described in more detail below.

The following approach is used for maintaining the vEB layout for centroid
trees in the dynamic COSB-tree: first, weight balance is relaxed from a
maximum weight difference of 2 between left and right subtrees, to a larger
constant $c$. Thus a node $v$ only becomes out of balance every
$\Omega(\textsc{weight(v)})$ insertions or deletions of descendants of v,
reducing the frequency of rebuilds. The nodes are stored in a PMA structure,
preserving order of the nodes using $\Theta(N)$space, apart from inserts and
deletes. On insert or delete, the structure is scanned in both directions
starting from the update point for sections of the array to rearrange based on
a "density" threshold of the section. Rearranging here refers to evenly
distributing the nodes across the section to make room for more nodes. To
ensure that updating pointers isn't too expensive, any node included in the
section must have its subtrees included. Specifically, the section is built up
from the the lowest subtree level that contains or contained the update point,
and neighboring subtrees are continuously added until either the density
threshold is met or the entire current subtree layout at this "level of
detail" is included. For example, suppose a node is inserted at the leaf of
some lower recursive subtree $L_{i}$. To continue building up this section,
$L_{i-1}$ or $L_{i+1}$ could be added. If all the lower recursive subtrees
$L_{1}$,$L_{2}$,...,$L_{x}$ have been added, the top level tree $U$ is added.
This corresponds to a subtree $L'_{j}$ one level above. Thus repeat this
process until the density threshold is met, and then rebalance the tree. This
approach depends on the fact that all the recursive subtrees at the same level
of detail asymptotically have the same size, which Bender et al. prove about
the hyperhyperfloor-based vEB layout described previously. This approach
results in an amortized update cost of $O(1+(\log{}N)/B)$ memory transfers per
update. In conjunction with the cost of rebuilds described above, this is only
an additive $O(\log{}N/B)$ factor, so the overall bound is amortized
$O(1+\log^{2}N/B)$ memory transfers.

Note that making the COSB-trees dynamic causes updates to successor and
predecessor pointers in the centroid tree to be expensive. Thus the notion of
ancestors is relaxed to the following (Lemma 10 in \cite{BenderFaKu06}): Let
$\alpha$ be the parent node of $\gamma$ in the compressed trie. Then either
$\alpha$ is a descendant of $\gamma$ or $\gamma$ is a descendant of $\alpha$.
The proof is as follows: As you recursively traverse the trie, both $\alpha$
and $\gamma$ are in the up trie or down trie, until at some point one of them
must be the centroid vertex. This is used to redefine what the predecessor
successor pointers refer to. The centroid may no loner point to the correct
predecessor or successor with respect to the relevant subtrie, but one of the
matched ancestors by construction must point to the correct predecessor or
successor. Thus for the dynamic case, when dealing with successors and
predecessors, $O(logN)$ ancestors' successor and predecessor pointers in
addition to the successor and predecessor pointers of the node in question
must be checked. However, this is an additive factor, so the same time bound
for update described above is still achieved.

\subsubsection{Modifying Front Compression to Preserve Locality}

This section provides a brief summary of how front compression is modified to
optimize compression and decoding for static and dynamic cache-oblivious
string B-trees. Consider a sequence of $j$ keys to store $k_{1}$, $k_{2}$,
..., $k_{j}$. Without compression, this requires $\sum_{j}||k_{j}||$ space in
memory. To compress these keys using front compression, let $\pi_{j}$ be the
longest common prefix between $k_{j}$ and $k_{j+1}$. Now instead we store the
keys in the following format:

\begin{center}
$k_{1}$,$||\pi_{2}||$,$\sigma_{2}$,$||\pi_{3}||$$\sigma_{3}$,...,$||\pi_{j}||$,$\sigma_{j}$
\end{center}

where $\sigma_{j}$ is the suffix of $k_{j}$ after removing the first $\pi_{j}$
bits. This saves almost $\sum_{j}||\pi_{j}||$ memory. However, since decoding
$k_{j}$ involves concatenating the first $\pi_{j}$ bits of $k_{j-1}$ to
$\sigma_{j}$, $k_{j-1}$ must be decoded. $k_{j-1}$ may depend on previous keys
as well, making $k_{j}$ expensive to decode.

The goal of modifying front compression is to maintain some of the space
benefits to compression while reducing the high decoding cost and keeping the
size of D small. Bender et al. provide a modification for static COSB-trees
that reduces the high decoding cost of front compression of D to
$O(1+||k||B)$, with size $O(\llangle{D}\rrangle{})$ after compression. The
modification works as follows: Let $c = 2 + \epsilon/2$, and let $k_{1}$ be
the first key added to the compression stream. For each subsequent key
$k_{i}$, if $k_{i}$ can be decoded with just the last $c||k_{i}||$ characters,
add $\pi_{i}$,$\sigma_{i}$ like before. Otherwise, add 0,$k_{i}$, which means
add $k_{i}$ to the stream in its uncompressed form. By construction, this
scheme guarantees that any key $k_{j}$ can be decoded in $O(1+||k||B)$ memory
transfers.%todo: add something about why this isO(<<D>>)?

However, the above modification falls short for anticipating insertions and
deletions. The previous modification relies on each key being decoded within
$c||k^{*}||$ elements of the compressed array, but this is no longer
guaranteed when new keys can be inserted to the left of $k$. To make the
locality-preserving front compression technique suitable for dynamic
COSB-trees, Bender et al. modify the algorithm to include an extra copied
prefix for each key. Consider the $3c||k||$ characters that would be to the
left of a new key$k^{*}$ after insertion. Let $k'$ be the key at the end of
this range. The idea is to prevent the insertion of $k^{*}$ from affecting any
non-copied keys just to the right of $k'$, since this could potentially
increase their decode cost. Compute the length $l$ of the largest common
prefix (lcp) between $k^{*}$ and $k'$, and compare $l$ to all the other lcp's
in this range. $l$ represents the minimum length for any lcp between $k^{*}$
and $k'$, since the keys are stored in lexicographic order. If there is a
copied prefix of length at least $l$ or a copied key within this range, insert
$k^{*}$ as usual. Inserting $k^{*}$ will have no negative effect in this
respect, since the decode length is effectively reset by the copied key or
prefix like before. However, if inserting $k^{*}$ causes the minimum length of
the lcp's in this range to increase, we need to fix the compression. To do
this, we update the lcp of the key directly to the right of $k^{*}$ and store
an extra copy of this lcp with $k^{*}$ itself. In this way, we can charge the
work of copying extra prefixes for a new key to the appropriate number of
elements before it. This number is determined by the length of the extra
copied prefix stored with the key.

\subsection{Operators}

%todo: figure out how to do range queries from this paper somehow...

The operations supported are dictated by the Dictionary Matching problem.
These operations are: \Search{} (synonymous with \Member{}), \Pred{}, \Succ{},
and \Range{}. The behavior of each operation for dynamic COSB-trees is
described below. \Insertkonly{} and \Delete{} are also supported but not
described here.

\Search{}: works as follows for a given key $k$. Compute the hash value for
$k$ and call it $\rho$, and traverse the centroid tree of the COSB-tree using
$\rho$ to obtain the successor $\rho_{s}$ and predecessor $\rho_{p}$ of $k$.
Index into the hashdata layer and perform a sequential scan across the PMA
structure from $\rho_{p}$ to $\rho_{s}$ to find $\rho$. On a match with value
$\rho'$, index into the keydata layer and decode $\rho'$ to obtain the
resulting key $k'$. If $k' = k$, return true. If $k' \ne k$ or $k$ was not
found during the sequential scan return false.

\Pred{},\Succ{}: works in the same way as \Search{} until the sequential scan
in the hashdata layer. When scanning the hashdata layer, look for $\rho_{p}$
or $\rho_{s}$, respectively, in place of $\rho$, and compare the appropriate
keys in the keydata layer.

\Range{}: The paper doesn't say, but presumably this is very similar to
\Search{}. Let $k_{b}$ be the beginning of the range and let $k_{e}$ be the
end of the range. Hash the range keys and identify the predecessor of
$\rho_{b}$ and the successor of $\rho_{e}$ using the centroid tree. Scan the
hashdata layer across from $\rho_{b,p}$ to $\rho_{e,s}$. Only start noting the
beginning of range when the first key with hash value after $\rho_{b}$, and
stop once we reach the first key with hash value after $\rho_{e}$. Decode and
return the set of keys within the given range in the keydata layer.

\subsection{Complexity}

\Search{} takes $O(1+\log_{B}N+||k||/B)$ memory transfers with high
probability (w.h.p.). Traversing the centroid tree requires $O(1+\log_{B}N)$
memory transfers, since the tree is built using$O(N/logN)$ keys and is
balanced. Once in the hashdata layer, we are already within $O(logN)$ keys of
$k$. Thus the local scan requires $O(1+(||k||+\log{}N)/B)$ memory transfers.
Obtaining $k$ from the keydata layer requires no more memory transfers than
the scan in the hashdata layer. The \Pred{} and \Succ{} operations are almost
identical in behavior to searches, and thus only have the added cost of having
to operate on $k'$, where $k'$ is the predecessor or successor, respectively.
These operations as a result take $O(1+\log_{B}N+||k||/B+||k'||/B)$ memory
transfers w.h.p.. Similarly, range queries take
$O(1+\log_{B}N+(||k||+||k'||+\llangle{}Q\rrangle{})/B)$ memory transfers,
where in this case $k'$ is the other end of the range and $Q$ is the set of
keys within the given range.

\Insertkonly{} and \Delete{} w.h.p. require $O(1+\log_{B}N+\log^{2}N||k||/B)$
memory transfers. As mentioned previously, due to rebuilds updates have
amortized cost of $O(1+(\log^{2}N)/B)$ memory transfers. The rest of the cost
for insertions and deletions is due to having to find the correct key to
delete or location for insert, which can be seen from the cost of \Search{}.

\section{Streaming B-trees}

\subsection{Description}

While B-trees achieve an optimal bound for the number of memory transfers for
search, they do not perform as well for bulk updates and insertions. The
buffered-repository tree (BRT) is an alternative to the B-tree that achieves an
improved amortized $O(\log{\frac{N}{B}})$ memory transfers for inserts compared
to the $O(\log_{B}{N})$ memory access of B-trees. In practice, log-structured
merge (LSM) trees \cite{ONeil96} are used for high-insert workloads.

This search vs. insertion trade-off also exists in the cache-oblivious model.
The original CO B-tree described in Section~\ref{sec:original} is designed to
optimize for search in this trade off. However, Bender et al. propose two data
structures that and achieve the opposite end: the shuttle tree and the
cache-oblivious lookahead array (COLA) \cite{BenderFaFi07}. These two data
structures are refereed to as \textit{streaming B-trees}, indicating how they
are optimized for bulk insertions at the cost of search.

\subsection{Shuttle Trees}

The shuttle tree is similar in layout and structure as the original CO B-tree.
It uses a strongly weight-balanced search tree described previously (see
Section~\ref{sec:structure}) that is laid out in a vEB (see
Figure~\ref{fig:veb}) layout into a PMA (see Section~\ref{sec:layout}). A
notable difference between the shuttle tree and the CO B-tree is that the CO
B-tree splits the tree at height $h/2$ when creating the vEB layout, while the
shuttle tree splits at height $F_k$, where $F_k$ is the $k$-th Fibonacci number
that is strictly less than the height of the tree. Thus if a tree's height
happens to be a Fibonacci number $F_k$, then the shuttle tree will choose to
split the tree at height $F_{k-1}$. The top and bottom ``halves'' produced by
this split are organized just like before and shown in Figure \ref{fig:veb}.

\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/buffers.pdf}
\end{center}

\caption{An illustration of how the shuttle tree determines the size of a node's linked list. The 
node $n2$ will have a linked list of size 2 because it is the leaf of two of the recursive subtrees. 
On the other hand $n1$ will only have a linked list of size 1.}
\label{fig:buffers}
\end{figure}

At a high level, the shuttle tree improves the efficiency of insertion by
adding buffers to each node in the tree. Shuttle tree nodes have a
linked list of buffers rather than just a single buffer like the BRT. Each of
these buffers is another shuttle tree. A shuttle tree node enforces two
constraints: the length of its linked list as well as the overall height (and
thus capacity) of each of the buffered shuttle trees contained within its
linked list are bounded, described below.

The length of a node's linked list is related to the node's height in the tree.
A node has a buffer for every time it is the leaf of a sub-tree that is split
when creating the vEB layout. This idea is illustrated in Figure
\ref{fig:buffers}.

The shuttle tree's node buffers have a maximum capacity that is enforced by
bounding its height. The size of the buffers increases going across the linked
list from left-to-right. Thus, a node's first buffer is smaller than its last
buffer. The maximum height of a buffer is given by $F_{H(k)} = F_{k-2 \lceil
\log_{\phi} k \rceil}$, where $\phi$ is the golden ratio and $k$ is chosen such
that $F_k$ = $\xi(h)$ and $\xi(x)$ is defined recursively: If $h$ is a
Fibonacci number then $\xi(h) = h$, otherwise $\xi(h) = \xi(h-f)$ where $f$ is
the largest Fibonacci number less than $h$. The maximum height of an individual
buffer is related to the size of the recursive subtree for which the node is a
leaf.  Since the recursive subtrees are split based on the largest Fibonacci
number less than $h$ (as described previously) it follows that the buffer sizes
are also a function of the Fibonacci numbers.

\subsubsection{Operators}

The shuttle tree supports \Search{}, \Insert{}, and \Scan{}. An explicit
discussion on \Delete{} is missing; it is assumed that it is similar 
to \Insert{}.

\Search{} and \Scan{} work mostly the same way as the CO B-tree. The only
difference is that a node's buffer must be searched before following a
pointer down the tree.

\Insert{} is also similar to the CO B-tree except elements are buffered as they
are inserted. \Insert{} starts by move down the tree from the root to leaves
and stopping when a node that has a buffer is located (note that because of the
nature of Fibonacci numbers that it is possible that a node does not have any
buffers). The element is inserted into the smallest buffer in the node's linked
list. If this insertion causes the buffer to overflow (e.g. the height of the
buffer exceeds its maximum), all of the elements from the buffer are moved into
the next largest buffer (and so on, recursively).  When the largest buffer
overflows all of its elements are taken and \textit{shuttled} down the tree.
Eventually a leaf node will be reached and inserting may cause the balance
constraints on the SWBST to become invalidated. The shuttle tree rebalances
itself just like any other SWBST: the leaf node is split into two and this
split cascades up the tree until balance is restored.
%% TODO: explain what it means to be "shuttled" down the tree

The shuttle tree also needs to take care of the buffers when a node is split.
When a node $u$ is split into two nodes $u_1$ and $u_2$, the shuttle tree
immediately allocates enough space for all of the $u_2$'s buffers ($u_1$ simply
reuses the space allocated for $u$'s buffers). $u$'s buffers are then scanned
over (by a single scan) and the elements are placed in the new buffers
accordingly.

The problem of maintaining room for inserts, bounding the costs of
merging and splitting, and bounding the costs of updating parent and child
pointers in the shuttle tree is achieved using the same PMA used with the CO
B-tree. This is described in detail in Section~\ref{sec:layout}.

\subsubsection{Complexity}

\Search{} uses $O(\log_{B}{N})$ memory transfers where $N = \Theta(c^{F_k})$ and
$c$ is the fanout. This bound is achieved by searching over the
top and bottom recursive subtrees at each level when traversing the tree. A
key trick is that when you visit a node you only search its largest buffer
(the last buffer in the linked list for that level). The smaller buffers will
be visited eventually as you recurse down the tree. Eventually you will get
down to a single subtree that can fit into a single blocks and thus can be
searched with one memory accesses. Since the fanout is proportional to the
block size it will take us $O(\log_B N)$ block accesses until we get down to a
subtree and thus, by induction, the worst case runtime is $O(\log_B N)$.

% Why is fanout propto block-size? doesn't this violate cache-obliviousness

% TODO - redo the insert section. Namely, include a discussion on how the
% shuttle tree maintains its structure dynamically. I mention the node
% splitting from pg 85 in the paper but do not include that in the analysis. I
% should basically mention that the cost of dynamically maintaining the
% shuttle tree O(logN / B) is not greater than the PMA log^2 N / B and thus
% the dynamic updates do not change the number of memory tranfers.

\Insert{} uses $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$
memory transfers. This improvement over the CO B-tree is achieved because the
shuttle tree's buffers are used to maintain locality of the elements long
enough to amortize the cost of moving them down the tree and possibly into
another memory block. The proof is based on the fact that when an element is
moved from one block to another it is due to either a leaf node splitting or a
node's buffer overflowing. Recall that when inserting a new element causes a
buffer to overflow, all of the the elements from that buffer move down the
tree, not just the single one that caused the overflow. Thus, the cost of
inserting an element is amortize by the number of elements in a node's largest
buffer. The number of elements in the largest buffer is
$B^{\Theta(1/(loglogB)^2)}$ leading to the $B^{\Theta(1/(loglogB)^2)}$ factor
in the big-Oh notation.

\Scan{} uses an amortized $O(L/B)$ memory transfers to scan $L$ consecutive
elements. This cost does not include the single \Search{} operation that is
required to find the first starting element of the range. Once the first
element's block is transferred, the shuttle tree proceeds by traversing the
tree and scanning in elements in order. Since the vEB layout enforces the
spatial locality of nearby nodes in the tree, the shuttle tree only has to read
in a new block about once every $B$ elements. Thus, each individual memory
access is is amortized to $1/B$ and the overall number of memory transfers is
$O(L/B)$.

The last thing that is worth mentioning is that the overall space requirements
of the shuttle tree is $O(N)$, since the majority of the buffers are really
small and thus do not impact the overall size of the tree.

\subsection{COLA}

The cache-oblivious lookahead array (COLA) is quite different from the other
cache-oblivious data structures discussed so far. Instead of an explicit tree
structure, the COLA is organized as a list of arrays or levels. Each array has
size $2^k$, where $k$ is the level of the array ($k = 0$ is the bottom level).
Furthermore, each array is either completely full or completely empty. As such,
the COLA is very similar to the binomial list structure described in
\cite{BentleySaxe}.

\subsubsection{Operators}

The operations supported by COLA are \Search{}, \Insert{}, and \Scan{}. Once
again, \Delete{} can be inferred from \Insert{}.

\Insert{} works by initially inserting an element into the first-level array.
This array has size $2^0 = 1$, thus the element is in an array on it own.
Whenever an element is inserted, check to see if there is another array of
the same size (this can be thought of as checking to see if there is another
array in the same level). If there is another array with the same size, the
COLA merges the two together in sorted order. This merging will produce a new
array that is twice the size and thus goes into the next level. The COLA
recursively merges until there are no two arrays of the same size.

\Search{} can naively be implemented by scanning each of the levels in the
COLA. Since everything is sorted the COLA can use a binary search to find the
element it is looking for. However, this is not very efficient. As an
improvement, COLA, as the name suggests, uses look ahead pointers to reduce the
number of memory accesses required. Every 8th element from the $(k+1)$-st level
exists in the $k$-th level array in its proper sorted location. This element
also contains a pointer to its real location in the $(k+1)$-st level. Thus this
pointer is referred to as \textit{real lookahead pointer}. Additionally, every
4th cell in $k$-th level array contains two pointers pointing to the next and
previous real lookahead pointers on this level. The next and previous pointers
are referred to as \textit{duplicate lookahead pointers}. To search a level the
COLA performs a binary search within an appropriate range, determine the range
in the next level using the duplicate lookahead pointers and then recursively
cascades up the tree.

\subsubsection{Complexity}

\Insert{} achieves an amortized $O(\log N / B)$ memory transfers. The proof is
based on the fact that an element will never be merged more than $O(\log{N})$
times. This is because at that point it will be in the last array.
Additionally, the first $\log{B}$ merges that a element participates in uses
$O(1)$ memory transfers since we assume that the relevant blocks are already in
memory. Thus, we only need to worry about memory transfers when we merge levels
that have size greater than $B$. If the size of the array is $k$, then the cost
of merging is $O(k/B)$, and thus the amortized cost of a merging a single
element is $1/B$. Since an element will only be merged $O(\log{N})$ time, it
follows that the overall amortized memory transfers to insert an element is
$O((\log N)/B)$.

\Search{} runs in $O(\log N)$ memory transfers. The proof uses induction to
show that the COLA never needs to search more than eight elements on each
level. Since there are $O(\log N))$ levels and since eight is a constant, it
follows that overall search uses $O(\log N)$ memory transfers. The fact that
only eight elements need to be searched per level is shown through induction.
The base case is trivial. The first three levels of the COLA each have at most
eight elements (note that these elements are either real elements or
lookahead pointers). The COLA can search these levels with no more
than eight member transfers. As the COLA follows lookahead pointers to the
higher levels, it will eventually get to a level where $k \geq 3$. Note that
the COLA does not ``enter'' this level at the start of the array; rather, its
arrived here by following the real lookahead pointers from the levels below
and thus is at some index $i$ that contains an element $e_i$ such that $e_i <
e$ and $e$ is the element being searched for. The goal at this point is to
find a new $i$ such that $e_i \leq e < e_{i+1}$. Once this $i$ is found the
duplicate lookahead pointers can be used to locate the \textit{next} and
\textit{previous} real lookahead pointers in the next level. The two real lookahead
pointers are used to restrict the search in the next level. Recall that at
each level in the COLA, every eighth element in a level is present in the
previous level. Thus, it follows that in the next level only 8 elements need
to be searched. Furthermore, since the current level was reach from a lower
level, it follows that only 8 elements need to be searched on this level to
find the appropriate $i$. Thus, overall, the COLA uses $O(8\log{N}) =
O(\log{N})$ memory transfers.

\subsubsection{Performance}

In experimental results, the COLA achieved significant improvements over
normal (cache-aware) B-trees, performing $790$ times faster. However this
improvement comes at the cost of sorted inserts and normal searches. Sorted
inserts are $3.1$ times slower and searches are $3.5$ times slower. Nevertheless,
these experimental results indicate that the COLA is a suitable data
structure when lots of random insertions are common.

\section{Concurrency}

\subsection{Motivation}

B-trees are often used as an indexing data structure for various database and
file systems. Thus, for a cache-oblivious B-tree data structure to be of
practical value, it must be able to handle concurrent access and modification
of its internal state. The straw-man solution is to simply ensure mutual
exclusion on the entire data structure and use the pre-existing data structures
without modification; this approach, however, yields poor performance.

Various concurrency schemes for traditional cache-aware B-trees have been
proposed \cite{BayerS77, LehmanY81}. We now survey a concurrency scheme for CO
B-trees, first presented in \cite{BenderFiGi05}. Bender develops two different
data structures to solve this problem: the first is an exponential CO B-tree,
the second is a CO B-tree based on a packed memory array (PMA). Bender also
presents a lock-free variant of the latter data structure.

\subsection{Description}

Before describing the data structures in more detail, we first consider
the concurrency model which is used throughout \cite{BenderFiGi05}.

\subsubsection{Concurrency Model}

The concurrency model starts with a system that has a total cache size of $M$
and block size of $B$. There are $P$ processors in the system, each with a
cache of size $\frac{M}{P}$. A particular block can reside in multiple caches
in \textit{shared} mode. However, if a processor wants to perform a write to a
block, it must acquire it in \textit{exclusive} mode, in which case that block
can only reside in a single cache. In this model, a request for a block costs
a single memory transfer.

For concurrency control, two primitives are used. The first is locks. The
second is \textit{load-linked/store-conditional} (LL/SC). The authors argue
that LL/SC provides better performance than using a \textit{compare-and-swap}
(CAS) primitive, but their techniques can be extended to use CAS.

\subsection{Exponential CO B-tree}

This data structure is based on the strongly weight-balanced exponential tree
described in Section~\ref{sec:structure}. A tuning parameter $1 < \alpha < 2$
is chosen for the data structure. Each node in the tree maintains both keys
which divide up its children, in addition to a \textit{right-link} pointer,
which contains both a pointer to a node's right sibling, plus the key value of
the minimum key in the node's right sibling. If no such sibling exists, a
sentinel $\infty$ pointer is used.

A node's layout in memory is as follows. Suppose a node has $k$ elements in
it. The first part of the node consists of a static CO search B-tree,
containing the $k$ elements. The tree is sized to be a complete $\lhyperceil k
\rhyperceil$-leaf tree. The second part of the node consists of a $\lhyperceil
k \rhyperceil$-size array containing the $k$ elements in sorted order. The leaf
nodes of the CO search tree point to the associated entries in the array.

\subsubsection{Operations}

The operations supported by the exponential CO B-tree are \Search{} and
\Insert{}. The authors do not discuss range-queries, and this is presumably
due to the concurrency issues in implementing a fully-consistent range scan
(weakly-consistent range scans could be implemented in the data structure with
very little modification), although interestingly enough the PMA variant of
B-tree (discussed in Section \ref{sec:pma}) does cover range scans.

\Search{} works as follows. For a non-leaf node, first check the
\textit{right-link} key. If the \textit{right-link} key is less than $k$, then
traverse the right link, and recurse. Otherwise, follow the appropriate child
pointer. When a leaf node is reached, only follow \textit{right-link}
pointers, until either $k$ has been located or some element greater than $k$
is reached. No locks are acquired during the entire search operation.

\Insert{} works as follows. First, \Search{} is used to locate the leaf node
where $k$ will be inserted into. A lock is acquired on that node, $k$ is
inserted, and the lock is released. Unlike a standard B-tree, there is no
maximum number of elements in a node; instead we split nodes
probabilistically. With probability ${1}/{2^{\alpha^h}}$, a node at height $h$
is split (leaf nodes have $h = 0$). A split works as follows. We reacquire a
lock on the node to be split. We split the node into $u$ and $u'$, where $u$
contains all the keys $< k$, and $u'$ contains the keys $\geq k$. We release
the lock, and recursively insert $k$ into the parent of the split node. It is
important to note that, since $u$ contains a prefix of the original node to be
split, that we reuse the original node in constructing $u$ (that way,
concurrent searches can still locate the keys in $u$).

%TODO: some section about linearizability?

\subsubsection{Complexity}

Sequential search takes $O(\log_B{N} + \log_{\alpha}{\log{B}})$ block transfers
(with high probability). The probabilistic bound comes from the probabilistic
key promotion property. The intuition behind the proof of this bound comes from
the argument that when the tree has height $h =
\Omega(\log_{\alpha}{\log{\log{N}}})$, then the total number of keys in the
tree is $O(2^{\alpha^h})$. The proof is then completed by summing over the
range of $h$ which is probabilistic $[0, O(\log_{\alpha}{\log{N}})]$, the
access time from some level $h$ to find the correct child.

Sequential insert takes $O(\log_B{N} + \log{\alpha}{\log{B}})$ block
transfers (in expectation). This proof is constructed in a similar fashion as
the search proof: consider inserting a tree at some height $h$. By noting that
the probability a particular node reaches height $h$ is
$2^{-(\alpha^h-1)/(\alpha-1)}$, a summation over all levels of the cost of
rebuilding a node yields the time-bound.

\subsection{Packed Memory Array CO B-tree}
\label{sec:pma}

The PMA CO B-tree is a two level data structure, containing a static CO B-tree
which indexes into a packed-memory array. Bender introduces a new
packed-memory array, called a \textit{one-way packed-memory array}, which he
argues is easier to manage concurrently.

\subsubsection{One-Way PMA Primer}

A one-way packed-memory array is a contiguous memory structure which has three
distinct regions: a leftmost region of empty space, an \textit{active region}
which contains the elements of the array, and a rightmost region of empty
space. The array only grows into the rightmost region, and shrinks from the
leftmost region (hence one-way). The PMA is of size $m = \Theta(N)$, subject
to $m$ being a power of $2$, where $N$ is the number of elements.

When the \textit{active region} becomes too large/small due to
inserts/deletes, the PMA is rebalanced. A new array is allocated, and the
elements are spread out evenly over the new array (to minimize element
density).

Each leaf in the static CO B-tree is a pointer into a $\Theta(\log N)$-size
region in the packed-memory array.

\subsubsection{Operations}

\Search{} works as follows. Search the static tree for $k$ until a leaf node
is reached. Follow the leaf node pointer into the PMA. Scan right in the PMA
until either $k$ is located, or some element $> k$ is reached. Do not acquire
any locks during the search.

\Insert{} works as follows. Use \Search{} to locate the $\Theta(\log
N)$-region of the PMA which (would) contain $k$. Lock the region, and scan
right to find the appropriate slot $s$ to place $k$ into. $s$ is defined such
that slot $s - 1$ contains the largest key smaller than $k$. If scanning
forward requires going into the next $\Theta(\log N)$-region, lock the next
region and release the current region (\textit{hand-over-hand} locking
technique). Once $s$ is located, either $s$ is free, in which case just insert
$k$, or $s$ is not free, in which case a \textit{rebalance} operation is
required.

To do a rebalance operation, we scan right (locking as necessary) until the
region from $s$ to the current position $s'$ is not \textit{too dense}. The
notion of density is defined in detail in \cite{BenderFiGi05}. The region $[s,
s']$ is then rebalanced as follows. Start from the rightmost element, and move
elements to the right (maintaining order) such that no suffix of $[s, s']$ is
too dense. This rebalancing creates room for $k$; we insert $k$ and then
release the locks.

\subsubsection{Complexity}

% TODO: can somebody check this time-bound? seems kind of bad

Sequential search time-bounds are not presented in the paper, but presumably
since search is a static CO tree lookup followed by a scan, the time bound is
$O(\log_{B}{N} + ({\log{N}})/{B} + 1)$.

Sequential insert is $O(\log_{B}{N} + (\log^2{N})/{B} + 1)$. The proof is based
off an accounting argument which places $\Theta((\log^2{m})/B)$ dollars in each
array slot. The proof is not presented in any detail in the paper, but rather
the authors argue that it is simply a slightly-modified version of the proof
presented in \cite{Katriel02}.

\subsubsection{Lock-free Variant}

The lock-free variant is based on the PMA CO tree, except the locks acquired
during the PMA modification are replaced by augmenting each element in the PMA
with a marker. A marker is used to denote the ongoing (concurrent) operation of
a particular cell. The authors assume that a marker is able to fit within a
single machine-word (so that the atomic operations are available and
performant).

In addition, four non-blocking primitives are introduced:
\begin{inparaenum}[(a)]
  \item \textit{move},
  \item \textit{cell-insertion},
  \item \textit{cell-deletion}, and
  \item \textit{read}
\end{inparaenum}. These primitives are self-explanatory, except they
all work in a non-blocking manner (they can fail at any point, and it is
up to the current process to restart the operation on failure).

The concurrency protocol works as follows. Any of the non-blocking primitives
are initiated by denoting the cell in question with the marker. If a process
ever encounters a cell with a marker, it drops what it is doing and first
helps to complete the operation. When the originating process gets
context-switched back in, any SCs will fail, and then the process can check to
see if the operation was already completed. The implementation of the
non-blocking primitives is spelled out in more detail in \cite{BenderFiGi05},
but the basic idea is using LL/SC to detect when concurrent operations are
on-going.

\subsubsection{Lock-free Variant Operations}

\Search{} is un-modified from the lock-based version.

\Insert{} is modified as follows. Cell $s$ is located as before. A
\textit{cell-insertion} operation is performed on $s$. If it succeeds, we are
done. If it fails, then a rebalance is necessary. Rebalance proceeds as
before, scanning right to find the region $[s, s']$. However, whenever a
non-empty cell is encountered, a load-link (LL) is performed on the cell. Once
the region is found, the rebalancing-to-the-right is performed as before.
However, a store-conditional (SC) is performed on the destination of each
cell. If any SC fails, then this means concurrent modification had occurred,
in which case the algorithm restarts from the beginning.

\section{Conclusion}
Fill this in.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
Fill this in.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}

\bibliography{citations}  % sigproc.bib is the name of the Bibliography in this case

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
%%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%% This next section command marks the start of
%% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The acm\_proc\_article-sp document class file itself is chock-full of succinct
%and helpful comments.  If you consider yourself a moderately
%experienced to expert user of \LaTeX, you may find reading
%it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}