### 2011-12-14 paper.tex
did my leg of the edits. Salman please skim for typos if you can. I found a few and may have missed/created some

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

% preprint + pagenumbering for page numbers
\documentclass[preprint]{style}
\pagenumbering{arabic}

\usepackage{paralist}
\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath}

% This gets rid of the block of whitespace on the first page
\makeatletter
\let\@copyrightspace\relax
\makeatother

\begin{document}

\title{Survey of Cache Oblivious B-trees}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Ahmad\\
\affaddr{MIT CSAIL}
\email{saahmad@mit.edu}
% 2nd. author
\alignauthor
Leilani Battle\\
\affaddr{MIT CSAIL}
\email{leibatt@mit.edu}
% 3rd. author
\alignauthor
Stephen Tu\\
\affaddr{MIT CSAIL}
\email{stephent@mit.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\date{7 December 2011}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\newcommand{\lhyperceil}{\lceil\lceil}
\newcommand{\rhyperceil}{\rceil\rceil}
\newcommand{\lhyperfloor}{\lfloor\lfloor}
\newcommand{\rhyperfloor}{\rfloor\rfloor}
\newcommand{\llangle}{\langle\langle}
\newcommand{\rrangle}{\rangle\rangle}

\newcommand{\Search}{\textsc{Search(k)}}
\newcommand{\Insert}{\textsc{Insert(k, v)}}
\newcommand{\Insertkonly}{\textsc{Insert(k)}}
\newcommand{\Scan}{\textsc{Scan(a, b)}}
\newcommand{\Delete}{\textsc{Delete(k)}}
\newcommand{\Succ}{\textsc{Succ(k)}}
\newcommand{\Pred}{\textsc{Pred(k)}}
\newcommand{\Member}{\textsc{Member(k)}}
\newcommand{\Range}{\textsc{Range-Query(k$_{i}$,k$_{j}$)}}

\maketitle
\begin{abstract}

This paper presents a survey of cache-oblivious B-trees (CO B-trees). The CO
B-tree was first introduced by Bender et al. \cite{BenderDemainColton}. It is
an external memory data structure, and thus it optimizes the number of memory transfers
performed rather than the number of CPU instructions executed. Unlike
other external memory data structures, like the widely used B-tree, the CO
B-tree does not have access to the parameters of the machine's memory
hierarchy, such as the block size/capacity at each level, or the relative
speeds. Hence, the CO B-tree is \textit{cache-oblivious}.  Despite this, the CO
B-tree achieves the same provably optimal search bound as the B-tree
and near optimal insertion and deletion bounds. In this paper we first describe
the original CO B-tree and explain how it achieves its performance bounds.
We then present and analyze three followup CO B-tree data structures that
improve upon weaknesses of the original: the \textit{cache-oblivious string
B-tree}, which introduces support for variable length keys; the
\textit{cache-oblivious streaming B-tree}, which improves the insertion bound
at the expense of the search bound; and the \textit{cache-oblivious
concurrent B-tree}, which introduces support for concurrency.

\end{abstract}

\section{Introduction}

\subsection{External Memory Algorithms}

Most external memory algorithms assume an idealized model of computation in which
memory is a flat byte-addressable array that can be read and modified in some
fixed time. However, in practice memory is typically
organized as a hierarchy consisting of fast on-chip caches (L1, L2, etc.),
slower DRAM, and even slower disks. When a program requests data, the data is
located in the memory hierarchy and typically carried up the chain to the
fastest memory layer (closest to the CPU), which is usually a cache. Unfortunately, this
cache has limited capacity, and once it is full any new data
requests necessitate some existing cached data to be evicted. Transferring data
from lower to higher levels in the memory hierarchy often takes an
exponentially long time, and access times between the different layers can differ
by many orders of magnitude. Therefore, the time to transfer data often
dominates the overall runtime performance of an algorithm.

External memory algorithms are wary of this fact, and focus on optimizing the
number of memory transfers performed instead of the number of CPU instructions
executed. They often try to exploit the fact that memory hierarchies transfer
data in blocks rather than individual machine word units. These
algorithms are designed to limit the number of memory transfers, as well as
preserve the locality of related data to amortize costly memory accesses.

\subsection{B-trees}

Because external memory algorithms are primarily concerned with data transfers,
many external memory problems can be reduced to finding
a suitable data structure that efficiently implements certain operations. A well
known external memory data structure is the B-tree \cite{BayerM72}, which is 
widely used in databases and file-systems. The B-tree
works by making every node in the tree the same size as a memory block in a
particular level of the hierarchy. The fanout of the tree
is also proportional to the block size.  The
B-tree incurs $O(\log_{B}{N})$\footnote{Throughout this paper, $\log$ with the
base omitted is base-$2$} memory transfers for \Search{}, \Insert{}, and
\Delete{}.

\subsection{DAM vs CO}

The B-tree has two theoretical shortcomings. First, the B-tree is analyzed in a
two-level memory model called the Disk Access Machine (DAM) \cite{Aggarwal}.
However, in practice, architectures typically have more than two levels in their memory hierarchies.
The argument for the DAM model is that in practice, only one
level in the memory hierarchy tends to dominate the performance, and thus a
two-level memory model is sufficient.  However, this means that B-trees may
incur an increasing number of memory transfers if the dominating level
in the hierarchy changes.

Additionally, the organization of a B-tree requires knowing the block size of
the memory layer in advance. B-trees are implemented with parameterized code
that needs to be modified for every level in the memory hierarchy. This tuning
can be very difficult and error prone, and may
even be impossible in certain heterogeneous computing environments.

The shortcomings of the DAM model gave rise to the cache-oblivious (CO) model
of computation \cite{Frigo99}. In the CO model an algorithm has no information
about the memory hierarchy, and knows nothing about the speed or the
block size of any level. Conceptually, the CO model allows algorithm designers
to reason about the simple two-level model like DAM, but also prove results for
any arbitrary multi-level memory hierarchy. This is because algorithms that
perform well in a CO model should, in theory, perform well regardless of
changes in block size, cache capacity, and cache speed. Consequently, CO
algorithms should gracefully scale to multi-level hierarchies as well.

\subsection{CO B-trees}

In \cite{BenderDemainColton}, Bender et al. introduce the cache-oblivious
B-tree (CO B-tree). The CO B-tree is similar to the normal B-tree except that
it does not require explicit information about the memory hierarchy.

The CO B-tree is able to achieve the optimal search bound of
$\Theta(\log_{B}{N})$ memory transfers, the optimal scan bound of $\Theta(N/B)$
memory transfers and near optimal insertion and deletion bound of 
$\Theta(\log_{B}{N} + \frac{\log^2{N}}{B})$. 
%The CO B-tree was a seminal data structure that gave rise to many others.

This paper provides a survey of the different types of CO B-trees and discusses
the various trade offs that they make. We start by analyzing the original CO
B-tree and explain how it works. This includes a detailed description of two data
structures integral to the CO B-trees performance: the
packed memory array (PMA) and strongly weight-balanced binary search tree
(SWBST). Following this
discussion we analyze three variants of the CO B-tree: the CO string B-tree (a CO B-tree with
variable length keys), the CO streaming B-tree (a CO B-tree that improves the
cost of insertion at the expense of search), and the CO concurrent B-tree (a CO
B-tree that allows multiple processes to efficiently access data concurrently).

\section{Background}

\subsection{Original CO B-trees}
\label{sec:original}

The CO B-tree optimizes the number of memory transfers by organizing its
elements as a strongly weight-balanced search tree (SWBST). The SWBST is then
laid out into a packed memory array (PMA) using a van Emde Boas (vEB) layout.
The SWBST and PMA play crucial roles in the other cache-oblivious data
structures that were developed after the CO B-tree, and thus are described at
length in the next two sections.

\subsection{Structure - SWBST}
\label{sec:structure}

\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/veb.pdf}
\end{center}

\caption{An illustration of the recursive van Emde Boas (vEB) layout}
\label{fig:veb}
\end{figure}

The ultimate goal of a CO B-tree is to reduce the number of memory transfers
while supporting \Search{}, \Insert{}, and \Delete{} operations. Thus, it is
not surprising that the CO B-tree maintains its data in a balanced search tree.
The search tree is laid out in memory according to a van Emde Boas (vEB) layout
\cite{veb1,veb2}: Let $h$
denote the height of a tree. The vEB layout splits the tree approximately at
height $h/2$. This splits the original tree into a \textit{top} subtree and
many \textit{bottom} subtrees.  Let $A$ denote the top subtree and
$B_1,...,B_k$ denote the bottom subtrees.  The vEB layout is defined
recursively using the order: $A,B_1,...,B_k$. This layout is also shown in
Figure~\ref{fig:veb}.

The vEB layout can be shown to use $O(\log_{B}{N})$ memory transfers for 
searches \cite{veb1, veb2}. Thus, using just the vEB layout, the CO
B-tree has already achieved the desired bound for search queries. However, it still
remains to be shown that the CO B-tree achieves the desired number of memory
transfers for dynamic inserts and deletes.

To support dynamic inserts/deletes, a \textit{strongly weight-balanced search
tree} (SWBST) is used \cite{swbst}.  Let the weight of a node $u$ be equal to the
total number of $u$'s descendants plus 1. A recursive definition for the weight of $u$ is the sum
of the weights of $u$'s children plus 1. Formally, $w(u) = 1 +
\sum_{v \in children} w(v)$.

Most weight-balanced search trees require that the heights of a node's left and
right subtrees differ by at most a constant factor. The SWBST enforces the
stronger property that a node at height $h$ in the tree (leaves have a height
of 1) has $\Theta(c^h)$ descendants, where $c$ is some constant ($c > 4$). This
property becomes important later on when laying the tree out in memory (see
Section~\ref{sec:layout}).

The definition of a SWBST given in \cite{swbst} provides this bound for the
weight of a node $u$:
\begin{equation}
  \label{eq:bounds}
  \frac{c^{h-1}}{2} \leq w(u) \leq 2c^{h-1}
\end{equation}

Let us now consider insertions. To insert an element, simply search through
the tree to find where the element would go. After inserting the element into some node $w$, 
check to see if any of $w$'s ancestors have become unbalanced by having a
weight greater than $2c^{h-1}$. If such an ancestor exists, it follows that
every node along the path between that ancestor and $w$ will also be unbalanced.
This requires splitting nodes to restore balance, starting from $w$
and moving up the tree.

When a node is split, we need to determine which children belong to which of
the two nodes. Dividing the children evenly is insufficient because it may
cause another violation. Instead, find the longest sequence of elements such
that the weight of the subsequence is $\lceil w(u)/2 \rceil$.  By doing this,
we maintain the bound that $\frac{c^{h-1}}{2} \leq w(u) \leq 2c^{h-1}$. Cascade
these changes up the tree until all of the ancestors are balanced.

Deletions are supported similarly, except we check if a node $u$ has a weight $w(u)
< \frac{c^{h-1}}{2}$. In this case, we merge two neighbors together and proceed
up the tree. However, the merging procedure may produce a new node that has a 
weight larger than the bounds
allow. If this happens, immediately split the node again, recursively. This can
be thought of as stealing nodes from a neighbor.

We have now sketched that the SWBST has the desired property that the number of
descendants of a node is $\Theta(c^h)$. We now describe the tree's memory layout.

\subsection{Layout - PMA}
\label{sec:layout}

Assuming the elements are now organized in a SWBST, we will show how to
of map this tree to memory and explain how it achieves the desired number
of memory transfers per operation. Do do this, we will use a packed memory 
array (PMA)\cite{BenderDemainColton}. There are three main costs
that the PMA addresses: \textit{inserts}, \textit{splits/merges}, and
\textit{pointer updates}.

First, the PMA needs to maintain free space for future inserts.
%Doing so may seem contradictory: maintaining the locality of elements while keeping space available for new inserts. 
To achieve a balance between maintaining locality 
and ensuring free space for new items, the PMA uses a technique
described in \cite{packedmemoryarray}.
The PMA performs an insertion into the SWBST using only $O(1 +
(\log^2{N}) / B)$ memory transfers. Note that this is the cost of updating the memory
layout of the SWBST in the PMA, not the cost of updating the SWBST%data structure
itself. At a high level, the PMA maintains
``windows'' into the array. When a window becomes too full or too empty, the
elements are spread out within the context of a larger window. Two parameters
are required: the window size and the threshold for triggering a spread out.
However, these parameters are not related to the block size of the memory level
and thus do not violate the CO model.

Second, the PMA needs to handle splits and merges caused by inserting into the
SWBST discussed in the previous section. The PMA is able to accomplish this
using $O(1 + (\log{N}) /B)$ memory transfers. The proof is based on the fact
that each node has $\Theta(c^h)$ descendants (recall that this is the key property
that SWBST provides, and is shown in Equation~\ref{eq:bounds}). Thus, merging 
or splitting a node will cause $\Theta(c^h)$ nodes to move, requiring 
$O(1 + c^h/B)$ memory transfers. It follows from the definition of SWBST in 
\cite{swbst} that when a node is rebalanced, it will require $\Theta(c^h)$ insertions 
or deletions until it is rebalanced again. Thus, the amortized cost of each insertion 
and/or deletion is $O(1/B)$. When a node is split, it may traverse all the way up the 
tree, and thus the split/merge bound is $O((\log N) / B)$ memory transfers.

Third, the PMA needs to update the parent and child pointers that
become invalidated as a result of the movement of the nodes. If the pointers are
all packed together in memory (e.g. within $B$ distance of each other) then
only $O(1)$ memory transfers are required. However, a problem arises if they 
are further apart. To address this issue, buffer nodes are introduced.  Buffer
nodes ``stretch'' the vEB layout so that nodes do not move far away from one
another. Using buffers, the PMA is able to achieve a pointer update cost
using $O(1+ \frac{\log{B}}{\sqrt{B}}\log^2{N})$ memory transfers.

Of these three terms, the cost of maintaining space for newly
inserted elements dominates - $O((\log^2{N})/B)$. Therefore, overall the CO B-tree
achieves insertions and deletions in $O(\log_B{N} + (\log^2{N})/B)$ memory
transfers. It still achieves the optimal $O(\log_B{N})$ memory transfers for
searches, matching the performance of static cache-aware B-trees.

\section{Streaming B-trees}

\subsection{Description}

While B-trees achieve an optimal bound in terms of memory transfers for
search, they do not perform as well for bulk updates and insertions. The
buffered-repository tree (BRT) is an alternative to the B-tree that achieves an
improved amortized $O(\log{\frac{N}{B}})$ memory transfers for inserts, compared
to the $O(\log_{B}{N})$ memory access of B-trees \cite{Buchsbaum}. In practice, log-structured
merge (LSM) trees \cite{ONeil96} are used for high-insert workloads.

This search vs. insertion trade-off also exists in the cache-oblivious model.
The original CO B-tree described in Section~\ref{sec:original} is designed to
optimize for search in this trade off. However, Bender et al. propose two data
structures that and achieve the opposite end: the shuttle tree and the
cache-oblivious lookahead array (COLA) \cite{BenderFaFi07}. These two data
structures are refereed to as \textit{streaming B-trees}, indicating how they
are optimized for bulk insertions at the cost of search.

\subsection{Shuttle Trees}

The shuttle tree is similar in layout and structure as the original CO B-tree.
It uses a strongly weight-balanced search tree described previously (see
Section~\ref{sec:structure}) that is laid out in a vEB (see
Figure~\ref{fig:veb}) layout into a PMA (see Section~\ref{sec:layout}). A
notable difference between the shuttle tree and the CO B-tree is that the CO
B-tree splits the tree at height $h/2$ when creating the vEB layout, while the
shuttle tree splits at height $F_k$, where $F_k$ is the $k$-th Fibonacci number
that is strictly less than the height of the tree. Thus if a tree's height
happens to be a Fibonacci number $F_k$, then the shuttle tree will choose to
split the tree at height $F_{k-1}$. The top and bottom ``halves'' produced by
this split are organized just like before and shown in Figure \ref{fig:veb}.

\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/buffers.pdf}
\end{center}

\caption{An illustration of how the shuttle tree determines the size of a node's linked list. The 
node $n2$ will have a linked list of size 2 because it is the leaf of two of the recursive subtrees. 
On the other hand $n1$ will only have a linked list of size 1.}
\label{fig:buffers}
\end{figure}

At a high level, the shuttle tree improves the efficiency of insertion by
adding buffers to each node in the tree. Shuttle tree nodes have a
linked list of buffers, compared to use of just a single buffer in the BRT. Each of
these buffers is another shuttle tree. A shuttle tree node bounds both the length 
of its linked list, and the overall height (and thus capacity) of each of the 
buffered shuttle trees contained within its linked list.

The length of a node's linked list is related to the node's height in the tree.
A node has a buffer for every time it is the leaf of a sub-tree that is split
when creating the vEB layout. This idea is illustrated in Figure
\ref{fig:buffers}.

The shuttle tree's node buffers have a maximum capacity that is enforced by
bounding its height. The buffers are arranged in increasing order of size from
left-to-right across the linked list. Thus, a node's first buffer is smaller than its last
buffer. The maximum height of a buffer is given by $F_{H(k)} = F_{k-2 \lceil
\log_{\phi} k \rceil}$, where $\phi$ is the golden ratio and $k$ is chosen such
that $F_k$ = $\xi(h)$ and $\xi(x)$ is defined recursively: If $h$ is a
Fibonacci number then $\xi(h) = h$, otherwise $\xi(h) = \xi(h-f)$ where $f$ is
the largest Fibonacci number less than $h$. The maximum height of an individual
buffer is related to the size of the recursive subtree for which the node is a
leaf.  Since the recursive subtrees are split based on the largest Fibonacci
number less than $h$ (as described previously), it follows that the buffer sizes
are also a function of the Fibonacci numbers.

\subsubsection{Operations}

The shuttle tree supports \Search{}, \Insert{}, and \Scan{}. An explicit
discussion on \Delete{} is missing; it is assumed that it is similar 
to \Insert{}.

\Search{} and \Scan{} work mostly the same as their counterparts in the original CO 
B-tree. The only difference is that a node's buffer must be searched before continuing
down the tree.

\Insert{} also behaves like insertions into the original CO B-tree, except that 
elements are buffered as they are inserted. \Insert{} starts by moving 
down the tree from the root to the leaves, and stopping when a node
that has a buffer is located. Note that because of the
nature of Fibonacci numbers, a node may not have any buffers. 
The element is inserted into the smallest buffer in the node's linked
list. If this insertion causes the chosen buffer to overflow (e.g. the height of the
buffer exceeds its maximum), all of the elements from the buffer are moved into
the next largest buffer, and so on recursively.  When the largest buffer
overflows, all of its elements are \textit{shuttled} down the tree.
Eventually, a leaf node is reached and the insertion may invalidate the balance
constraints on the SWBST. The shuttle tree rebalances
itself as described previously in Section~\ref{sec:structure}: the leaf node is 
split into two and this split cascades up the tree until balance is restored.
%% TODO: explain what it means to be "shuttled" down the tree

The shuttle tree also needs to manage buffers when a node is split.
When a node $u$ is split into two nodes $u_1$ and $u_2$, the shuttle tree
immediately allocates enough space for all of $u_2$'s buffers. $u_1$
reuses the space allocated for $u$'s buffers. Then a single scan is performed
over $u$'s buffers, and the elements are placed in the new buffers
accordingly.

The PMA structure described in Section~\ref{sec:layout} is used to: maintain
room for inserts, bound the costs of merging and splitting, and bound the 
costs of updating parent and child pointers in the shuttle tree

\subsubsection{Complexity}

\Search{} uses $O(\log_{B}{N})$ memory transfers, where $N = \Theta(c^{F_k})$ and
$c$ is the fanout. This bound is achieved by searching over the
top and bottom recursive subtrees at each level when traversing the tree. A
key observation is that when you visit a node you only search its largest buffer
(the last buffer in the linked list for that level). The smaller buffers will eventually
be visited as you recurse down the tree. Eventually, you will traverse
down to a single subtree that fits in a single block, and thus can be
searched with one memory transfer. Since the fanout is proportional to the
block size, it will only take $O(\log_B N)$ block accesses to reach this
subtree. Thus, by induction, the worst case runtime is $O(\log_B N)$.

% Why is fanout propto block-size? doesn't this violate cache-obliviousness

% TODO - redo the insert section. Namely, include a discussion on how the
% shuttle tree maintains its structure dynamically. I mention the node
% splitting from pg 85 in the paper but do not include that in the analysis. I
% should basically mention that the cost of dynamically maintaining the
% shuttle tree O(logN / B) is not greater than the PMA log^2 N / B and thus
% the dynamic updates do not change the number of memory tranfers.

\Insert{} uses $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$
memory transfers. This improvement over the CO B-tree is achieved using
the shuttle tree's buffers to maintain locality of the elements long
enough to amortize the cost of moving them down the tree and possibly into
another memory block. The proof is based on the fact that when an element is
moved from one block to another, it is due to either a leaf node splitting or a
node's buffer overflowing. Recall that when inserting a new element causes a
buffer to overflow, all of the the elements from that buffer move down the
tree, not just the single element that caused the overflow. Thus, the cost of
insertion is amortized by the number of elements in a node's largest
buffer. The number of elements in the largest buffer is
$B^{\Theta(1/(loglogB)^2)}$ leading to the $B^{\Theta(1/(loglogB)^2)}$ factor
in the big-O notation.

\Scan{} uses an amortized $O(L/B)$ memory transfers to scan $L$ consecutive
elements. This cost does not include the single \Search{} operation that is
required to find the starting element of the range. Once the first
element's block is transferred, the shuttle tree proceeds by traversing the
tree and scanning the elements in order. Since the vEB layout enforces
spatial locality of nearby nodes in the tree, the shuttle tree only has to read
in a new block roughly once every $B$ elements. Thus, each individual memory
access is is amortized to $1/B$ memory transfers, and the total number of 
memory transfers is $O(L/B)$.

Note that the shuttle tree uses $O(N)$ space, since 
the majority of the buffers are very small and thus do not impact the 
overall size of the tree.

\subsection{COLA}

The cache-oblivious lookahead array (COLA) is notably different from the other
CO data structures discussed thus far. Instead of an explicit tree
structure, the COLA is organized as a list of arrays or levels. Each array has
size $2^k$, where $k$ is the level of the array ($k = 0$ is the bottom level).
Furthermore, each array is either completely full or completely empty. As such,
the COLA is very similar to the binomial list structure described in
\cite{BentleySaxe}.

\subsubsection{Operations}

The operations supported by COLA are \Search{}, \Insert{}, and \Scan{}. Once
again, \Delete{} can be inferred from \Insert{}.

\Insert{} works by initially inserting an element into the first-level array.
This array has size $2^0 = 1$, thus the element is in its own array.
Each time an element is inserted, check to see if there is another array of
the same size (this can be thought of as checking to see if there is another
array in the same level). If there is another array with the same size, the
COLA merges the two together in sorted order. This merging will produce a new
array that is twice the size and thus is moved into the next level. The COLA
recursively merges until there are no two arrays of the same size.

\Search{} can naively be implemented by scanning each of the levels in the
COLA. Since all the arrays are sorted the COLA can use a binary search to find the
element it is looking for. However, this is not very efficient. As an
improvement COLA, as the name suggests, uses look ahead pointers to reduce the
number of required memory accesses. Every 8th element from the $(k+1)$-st level
exists in the $k$-th level array in its proper sorted location. This element
also contains a pointer to its real location in the $(k+1)$-st level. Call this
pointer a \textit{real lookahead pointer}. Additionally, every
4th cell in a $k$-th level array contains two pointers pointing to the next and
previous real lookahead pointers on this level. The next and previous pointers
are referred to as \textit{duplicate lookahead pointers}. To search a level the
COLA performs a binary search within an appropriate range, determines the range
in the next level using the duplicate lookahead pointers, and then recursively
cascades up the tree.

\subsubsection{Complexity}

\Insert{} achieves an amortized $O(\log N / B)$ memory transfers. The proof is
based on the fact that an element will never be merged more than $O(\log{N})$
times, since at that point it will be in the last array.
Additionally, the first $\log{B}$ merges that a element participates in uses
$O(1)$ memory transfers since we assume that the relevant blocks are already in
memory. Thus, we only need to worry about memory transfers when we merge levels
that have size greater than $B$. If the size of the array is $k$, then the cost
of merging is $O(k/B)$, and thus the amortized cost of a merging a single
element is $1/B$. Thus it follows from our earlier assumption that the 
overall amortized number of memory transfers needed to insert an element is 
$O((\log N)/B)$.

\Search{} runs in $O(\log N)$ memory transfers. The proof shows that
the COLA never needs to search more than 8 elements on each
level. Since there are $O(\log N))$ levels and 8 is a constant, it
follows that overall search uses $O(\log N)$ memory transfers. The fact that
only 8 elements need to be searched per level is shown through induction.
The base case is trivial. The first three levels of the COLA each have at most
8 elements (note that these elements are either real elements or
lookahead pointers). The COLA can search these levels with no more
than 8 member transfers. As the COLA follows lookahead pointers to the
higher levels, it will eventually reach a level where $k \geq 3$. Note that
the COLA does not ``enter'' this level at the start of the array. It
arrives there by following the real lookahead pointers from the levels below,
and thus is located at some index $i$ that contains an element $e_i$ such that $e_i <
e$ and $e$ is the element being searched for. Now we want to
find a new $i$ such that $e_i \leq e < e_{i+1}$. Once this $i$ is found, the
duplicate lookahead pointers are used to locate the \textit{next} and
\textit{previous} real lookahead pointers in the next level. The two real lookahead
pointers are used to restrict the search in the next level. Recall that at
each level in the COLA, every 8 element in a level is present in the
previous level. Thus, it follows that in the next level only 8 elements need
to be searched. Furthermore, since the current level was reached from a lower
level, it follows that only 8 elements need to be searched on this level to
find the appropriate $i$. Therefore, overall, the COLA uses $O(8\log{N}) =
O(\log{N})$ memory transfers.

\subsubsection{Performance}

In experimental results, the COLA achieved significant improvements over
normal (cache-aware) B-trees, performing $790$ times faster. However this
improvement comes at the cost of sorted inserts and normal searches. Sorted
inserts are $3.1$ times slower and searches are $3.5$ times slower. Nevertheless,
these experimental results indicate that the COLA is a suitable data
structure when lots of random insertions are common.

\section{Variable Length Strings}
\begin{figure}

\begin{center}
	\includegraphics[width=\columnwidth]{figures/COSB-treegeneralstructurediagram.pdf}
\end{center}

\caption{An illustration of the dynamic COSB-tree structure}
\label{fig:string_structure}
\end{figure}
\subsection{Motivation}

In order for B-trees to remain versatile, they must be able to accommodate
various forms of data. Bender et al. argue the following weaknesses in
existing B-tree designs: 
\begin{inparaenum}[(a)]
  \item traditional B-trees have suboptimal performance when variable length
  strings are used as keys \cite{BenderDemainColton},
  \item string B-trees \cite{Ferragina98} handle variable-length strings well
  but do not use compression, 
  \item front compression \cite{BayerUn77,ClarkSaSt69,Comer79,Wagner73} is used
  in many B-trees, but is handled sub-optimally \cite{BenderFaKu06}. 
\end{inparaenum}

In \cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache
Oblivious String B-tree (COSB-tree), which works more efficiently
for variable length string keys. The COSB-tree uses a modified front-compression
scheme, which uses less space and improves compressed data locality.

\subsubsection{General COSB-tree Structure}

The dynamic COSB-tree consists of three layers: a weight-balanced centroid tree stored in
a packed-memory array (PMA, see Section~\ref{sec:layout}); a PMA of hash data for the keys, known as the 
\emph{hashdata} layer; and a PMA of compressed values for the original key,
known as the 
\emph{keydata} layer (see Figure~\ref{fig:string_structure}). Each leaf of the centroid tree contains a pointer into
the hashdata array. The centroid tree contains only hash values, and is designed such that
each leaf points to the respective \emph{neighborhood} of keys being
searched for. After traversing the centroid tree and following the pointer into
the hashdata array layer, we are within $O(\log{}N)$ keys of $k$. Then a
sequential scan is performed in the hashdata array layer to find $k$. Each
entry in the hashdata array contains only the hash value of the corresponding key,
and a pointer to the compressed key value in the keydata layer. The keydata layer
is consulted to resolve hash collisions in the hashdata layer, and to obtain the original value of the key.

In the following sections, centroid trees are described, and a summary of
how to modify them to preserve balance and structure in the dynamic case is
provided. Two iterations on forward compression are also summarized to show
how to incrementally add support for variable length strings as keys for
dynamic COSB-trees.

\subsubsection{Centroid Trees}
\label{sec:centroid_trees}
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidtreeexample.pdf}
\end{center}

\caption{A centroid tree example from \cite{BenderFaKu06}. Dotted lines
represent centroid-rooted up tries, where the parent is not a prefix of the
child. Solid lines represent centroid-rooted down tries, where the parent is a
prefix of the child.}
\label{fig:centroid}
\end{figure}

COSB-trees are a solution to the \textit{dictionary matching problem}.
Dictionary matching, or string matching, refers to matching strings using
prefix search. Let $D$ be a dictionary of $N$ keys. To perform faster searches,
the keys in $D$ are preprocessed using hashing and divide-and-conquer via
centroid trees. Figure~\ref{fig:centroid} depicts an example centroid tree from
\cite{BenderFaKu06}.

Consider the compacted\footnote{Compacted means all non-branching paths are
reduced to edges.} trie $T$ constructed over $D$ for string matching. There exists a
vertex $\rho$ in $T$ such that $\rho$ has between N/3 and 2N/3 descendants,
inclusive. Call $\rho$ the \emph{centroid vertex} of $T$. When searching for a
key $k$ in the compacted trie, you traverse the trie as follows: for each node
$y$ that you visit, if $y$ is a prefix of $x$, or $y$ matches $x$, continue on
into the trie rooted at $y$ or \emph{down trie} of $y$. Otherwise traverse the
trie resulting from excluding $y$ and its subtree, and call this the \emph{up trie} of
$y$. The \emph{centroid tree} is obtained by rooting the tree at the centroid vertex
$\rho$, and making its children recursively defined centroid trees of its up and
down tries.

The centroid tree in terms of structure is essentially a weight-balanced binary tree (see
Section~\ref{sec:structure}), and is laid out in memory using a modified vEB
layout (see Figure~\ref{fig:veb}). However, instead of using height $h/2$ as
described previously, the \emph{hyperhyperfloors} of the weights of nodes are
used as a guide for cutting the tree into equal-size subtrees for the vEB layout. The
\emph{hyperhyperfloor} of x refers to rounding x down to the nearest power of a 
power of 2.

In addition, a different approach is used to rebalance the centroid tree: put all the
elements in DFS or Euler tour order for the compacted trie. Given this order,
scan the trie to identify the centroid vertex. Then partition the trie into its
respective up and down tries, and recurse on the up and down tries. The Euler
tour order can be found in $O(\log{N})$ scans, since the trie is only
$O(\log{N})$ deep. Thus these updates have amortized cost of
$O(1+(\log^{2}{N})/B)$ memory transfers.

Similar to the original CO B-tree, the centroid tree is stored in a PMA
(see Section~\ref{sec:layout}). However, the centroid tree PMA is more 
flexible with a modified updating scheme. The changes made to updating 
"windows" into the array are as follows (see Section~\ref{sec:layout} 
for more on "window" updates): on insert or 
delete, the area around the update point is scanned for sections
or windows of the array to rearrange. If a window of a
certain \textit{density} threshold is found (as described previously in~\ref{sec:layout}), the 
nodes are evenly distributed across the
window to make room for more nodes. To ensure that updating pointers is not too
expensive, any node included in the window must have all of its subtrees
included. The subtrees are dictated by the vEB layout, and subtrees are chosen
from the bottom up. The smallest neighboring subtrees (single nodes) in the vEB
layout are added first, and larger subtrees are included when all of their own
subtrees have already been added to the section. top subtrees are only added after all
bottom subtrees are added (in terms of the vEB layout). Figure
\ref{fig:centroid_relayout} gives an example of how subtrees are chosen for
redistribution of nodes.

This approach results in an amortized update cost of $O(1+(\log{N})/B)$ memory
transfers per update. In conjunction with the cost of rebalancing described above,
this is only an additive $O((\log{N})/B)$ factor, so the overall bound is
amortized $O(1+(\log^{2}N)/B)$ memory transfers.
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidrelayoutexample.pdf}
\end{center}

\caption{A centroid layout example. Assume the smallest subtrees in this
example only contain a single node. Suppose a node was inserted into a leaf
position in the bottom left of the tree. The numbers represent the order in
which each subtree is added to the window to be rearranged.}
\label{fig:centroid_relayout}
\end{figure}

\subsubsection{Modifying Front Compression}

This section provides a brief summary of front compression, and how it is modified to
optimize compression and decoding for dynamic COSB-trees. This approach 
increases the locality of the compressed data.

First, front compression is described: consider a sequence
of $j$ keys to store $k_{1},k_{2},...,k_{j}$. Without compression, this
requires $\sum_{j}||k_{j}||$ space in memory. To compress these keys using
standard front compression, let $\pi_{j}$ be the longest common prefix between
$k_{j}$ and $k_{j+1}$. Now instead we store the keys in the following format:
\begin{equation}
  k_{1}||\pi_{2}||\sigma_{2}||\pi_{3}||\sigma_{3}||...||\pi_{j}||\sigma_{j}
\end{equation} 
where $\sigma_{j}$ is the suffix of $k_{j}$ after removing the first $\pi_{j}$
bits. This saves almost $\sum_{j}||\pi_{j}||$ memory. However, since decoding
$k_{j}$ involves concatenating the first $\pi_{j}$ bits of $k_{j-1}$ to
$\sigma_{j}$, $k_{j-1}$ must be decoded. $k_{j-1}$ may depend on previous keys
as well, making $k_{j}$ expensive to decode.

Bender et al. provide a modification that reduces the high decoding cost of
front compressing $D$ to $O(1+||k||B)$ memory transfers, while 
maintaining a compressed size of
$O(\llangle{D}\rrangle{})$\footnote{$\llangle{D}\rrangle{}$ is shorthand for
the size of $D$ after front compression} after compression. The modification
works as follows: Let $c = 2 + \epsilon/2$, and let $k_{1}$ be the first key
added to the compression stream. For each subsequent key $k_{i}$, if $k_{i}$
can be decoded with just the last $c||k_{i}||$ characters in the compression stream, add
$\pi_{i}$, $\sigma_{i}$ as described above for standard front compression.
Otherwise, add $k_{i}$ to the stream
without compressing it. By construction, this scheme guarantees that any key
$k_{j}$ can be decoded in $O(1+||k||B)$ memory transfers.

However, the above modification falls short for anticipating insertions and
deletions. The previous modification relies on each key being decoded within
$c||k||$ elements of the compressed array, but this is no longer guaranteed
when new keys can be inserted to the left of $k$. To make the
locality-preserving front compression technique suitable for dynamic
COSB-trees, Bender et al. modify the algorithm to include extra copied
prefixes. Suppose a key $k$ is being inserted. If keys further down in the
array will be negatively impacted, the longest common prefix between $k$ and
the key right after $k$ is copied to $k$ along with $k$'s original compression
data. The idea behind adding extra copied prefixes is to maintain $O(||k||)$
decode distances around the area where the new key is being inserted. The
characters preceding $k$ are charged for the copied prefix, accounting for the
work to add the copied prefix to $k$.

\subsection{Operations}

The operations supported by the COSB-tree are: \Search{} (synonymous with \Member{}), \Pred{}, \Succ{},
and \Scan{}. The behavior of each operation for dynamic COSB-trees is
described below. \Insertkonly{} and \Delete{} are also supported but not
described here.

\Search{} works as follows for a given key $k$. Compute the hash value for
$k$ and call it $\rho$. Traverse the centroid tree of the COSB-tree using
$\rho$ to obtain the successor $\rho_{s}$ and predecessor $\rho_{p}$ of $k$.
Index into the hashdata layer with the resulting pointer from the centroid tree traversal,
and perform a sequential scan across the PMA
from $\rho_{p}$ to $\rho_{s}$ to find $\rho$. On a match with hash value
$\rho'$, index into the keydata layer and decode $\rho'$ to obtain the
resulting key $k'$. If $k' = k$, return true. If $k' \ne k$ or $k$ was not
found during the sequential scan return false.

\Pred{}, \Succ{} work in the same way as \Search{} until the sequential scan in
the hashdata layer. When scanning the hashdata layer, look for $\rho_{p}$ or
$\rho_{s}$, respectively, in place of $\rho$, and compare the appropriate keys
in the keydata layer.

\Scan{}: This operation is not explicitly described in \cite{BenderFaKu06}, but
presumably this is very similar to \Search{}. Let $k_{b}$ be the beginning of
the range and let $k_{e}$ be the end of the range. Hash the range keys and
identify the predecessor of $\rho_{b}$ and the successor of $\rho_{e}$ using
the centroid tree. Scan the hashdata layer across from $\rho_{b,p}$ to
$\rho_{e,s}$. Only start noting the beginning of the range when the first key with
hash value after $\rho_{b}$, and stop once we reach the first key with hash
value after $\rho_{e}$. Decode and return the set of keys within the given
range in the keydata layer.

\subsection{Complexity}
The keys are hashed and hash values are compared in place of the original key
values to speed up comparisons in the centroid tree and hashdata layer. To
avoid mismatches due to hash collisions, the original key values are compared
only once at the bottom layer, resulting an additive $O(||k||/B)$ memory
transfers in each of the operators for each key $k$ that needs to be compared.

\Search{} takes $O(1+\log_{B}N+||k||/B)$ memory transfers with high probability
(w.h.p.).  The centroid tree is stored in a PMA with a dynamic CO layout. This
layout reduces the number of memory transfers required when traversing the tree
to $O(\log_{B}{N})$.  Once in the hashdata layer, we are already within $O(\log{N})$
keys of $k$, which require $O((\log{N})/B)$ memory transfers to scan. Decoding
and comparing the original key values requires $O(1+||k||/B)$ memory transfers,
due to the modified front compression scheme.

The \Pred{} and \Succ{} operations are similar in behavior to searches, and
thus only have the added cost of having to operate on $k'$, where $k'$ is the
predecessor or successor, respectively. These operations as a result take
$O(1+\log_{B}N+||k||/B+||k'||/B)$ memory transfers (w.h.p.). Similarly, range
queries take $O(1+\log_{B}N+(||k||+||k'||+\llangle{}Q\rrangle{})/B)$ memory
transfers, where in this case $k'$ is the other end of the range and $Q$ is the
set of keys in $D$ that are within the given range.

\Insertkonly{} and \Delete{} require $O(1+\log_{B}N+\log^{2}N||k||/B)$
memory transfers (w.h.p.). As mentioned previously in Section~\ref{sec:centroid_trees}, 
due to rebalancing updates have amortized cost of $O(1+(\log^{2}N)/B)$ memory transfers.
The rest of the cost for insertions and deletions is due to having to find the correct key to
delete or location for insert, which can be seen from the cost of \Search{}.

\subsection{Experimental Results}
Bender et al. perform two experiments in \cite{BenderFaKu06}, in which a
regular B-tree implementation tuned to various block sizes is compared to a
static COSB-tree implementation. The experiment involved measuring the amount
of time required to perform:
\begin{inparaenum}[(a)]
  \item 440,000 and 450,000 inserts of random values, 
  \item a range query of all values, and 
  \item 1000 random searches.
\end{inparaenum} The results of the
experiment show that COSB-trees outperformed B-trees most of the time for
random searches, range queries and random inserts. However, some point between
440,000 and 450,000 inserts the COSB-tree had to rebalance the entire tree,
causing it to fall behind in this measurement. However, Bender et al. argue
that the experiment was biased towards the B-tree implementation, since the
trees are ``young'', and thus still well organized. The COSB-tree
implementation is also compared to Berkeley DB \cite{BerkeleyDB}, and achieves
performance comparable to Berkeley DB with tuned parameters. The results seem
to show that there is a window in which large and small-block size B-trees
perform poorly, and that COSB-trees may be a reasonable alternative to improve
performance.

\section{Concurrency}

\subsection{Motivation}

B-trees are often used as an indexing data structure for various database and
file systems. Thus, for a cache-oblivious B-tree data structure to be of
practical value, it must be able to handle concurrent access and modification
of its internal state. The straw-man solution is to simply ensure mutual
exclusion on the entire data structure and use the pre-existing data structures
without modification. However, this approach yields poor performance.

Various concurrency schemes for traditional cache-aware B-trees have been
proposed \cite{BayerS77, LehmanY81}. We now survey a concurrency scheme for CO
B-trees, first presented in \cite{BenderFiGi05}. Bender et al. develops two different
data structures to solve this problem: first an exponential CO B-tree,
and second a CO B-tree based on a PMA (see Section~\ref{sec:layout} for more on PMA's). Bender et al. also
present a lock-free variant of the latter data structure.

\subsection{Description}

Before describing the data structures in more detail, we first consider
the concurrency model used throughout \cite{BenderFiGi05}.

\subsubsection{Concurrency Model}

The concurrency model starts with a system that has a total cache size of $M$
and block size of $B$. There are $P$ processors in the system, each with a
cache of size $\frac{M}{P}$. A particular block can reside in multiple caches
in \textit{shared} mode. However, if a processor wants to perform a write to a
block, it must acquire it in \textit{exclusive} mode, in which case that block
can only reside in a single cache. In this model, a request for a block costs
a single memory transfer.

For concurrency control, two primitives are used. The first is locks. The
second is \textit{load-linked/store-conditional} (LL/SC). The authors argue
that LL/SC provides better performance than using a \textit{compare-and-swap}
(CAS) primitive, but their techniques can be extended to use CAS.

\subsection{Exponential CO B-tree}

This data structure is based on the strongly weight-balanced exponential tree
described in Section~\ref{sec:structure}. A tuning parameter $1 < \alpha < 2$
is chosen for the data structure. Each node in the tree maintains the keys
which divide up its children, in addition to a \textit{right-link} pointer. The right-link
pointer contains a pointer to the node's right sibling, and the key value of
the minimum key in the node's right sibling. If no such sibling exists, a
sentinel $\infty$ pointer is used.

A node's layout in memory is as follows. Suppose a node has $k$ elements in
it. The first part of the node consists of a static CO search B-tree,
containing the $k$ elements. The tree is sized to be a complete $\lhyperceil k
\rhyperceil$-leaf tree. The second part of the node consists of a $\lhyperceil
k \rhyperceil$-size array containing the $k$ elements in sorted order. The leaf
nodes of the CO search tree point to the associated entries in the array.

\subsubsection{Operations}

The operations supported by the exponential CO B-tree are \Search{} and
\Insert{}. The authors do not discuss range-queries, and this is presumably
due to the concurrency issues in implementing a fully-consistent range scan
(weakly-consistent range scans could be implemented in the data structure with
very little modification). However, the PMA variant of CO
B-tree (discussed in Section \ref{sec:pma}) does cover range scans.

\Search{} works as follows. For a non-leaf node, first check the
\textit{right-link} key. If the \textit{right-link} key is less than $k$, then
traverse the right link, and recurse. Otherwise, follow the appropriate child
pointer. When a leaf node is reached, only follow \textit{right-link}
pointers, until either $k$ has been located or some element greater than $k$
is reached. No locks are acquired during the entire search operation.

\Insert{} works as follows. First, \Search{} is used to locate the leaf node
that $k$ will be inserted into. A lock is acquired on that node, $k$ is
inserted, and the lock is released. Unlike a standard B-tree, there is no
maximum number of elements in a node. Instead, we split nodes
probabilistically: with probability ${1}/{2^{\alpha^h}}$, a node at height $h$
is split (leaf nodes have $h = 0$). To perform a split, we first reacquire a
lock on the node to be split. We split the node into $u$ and $u'$, where $u$
contains all the keys $< k$, and $u'$ contains the keys $\geq k$. We then release
the lock, and recursively insert $k$ into the parent of the split node. It is
important to note that, since $u$ contains a prefix of the original node to be
split, we reuse the original node in constructing $u$. Doing so ensures that
concurrent searches can still locate the keys in $u$.

%TODO: some section about linearizability?

\subsubsection{Complexity}

Sequential search takes $O(\log_B{N} + \log_{\alpha}{\log{B}})$ block transfers
(w. h. p.). The probabilistic bound comes from the probabilistic
key promotion property. The intuition behind the proof of this bound comes from
the argument that when the tree has height $h =
\Omega(\log_{\alpha}{\log{\log{N}}})$, the total number of keys in the
tree is $O(2^{\alpha^h})$. The proof is then completed by summing over the
range of $h$ which is probabilistic $[0, O(\log_{\alpha}{\log{N}})]$, which is the
the number of memory transfers made, starting from some level $h$,
to find the correct child.

Sequential insert takes $O(\log_B{N} + \log{\alpha}{\log{B}})$ block
transfers (in expectation). This proof is similar to the search proof: 
consider inserting a tree at some height $h$. By noting that
the probability that a particular node reaches height $h$ is
$2^{-(\alpha^h-1)/(\alpha-1)}$, a summation over all levels of the cost of
rebuilding a node yields the desired bound.

\subsection{Packed Memory Array CO B-tree}
\label{sec:pma}

The packed-memory array (PMA) CO B-tree is a two level data structure, containing a static CO B-tree
which indexes into a PMA (see Section~\ref{sec:layout} for more on PMA's). 
Bender et al. introduce a new PMA, 
called a \textit{one-way packed-memory array}, which he
argues is easier to manage concurrently.

\subsubsection{One-Way PMA Primer}

A one-way PMA is a contiguous memory structure which has three
distinct regions: a leftmost region of empty space, an \textit{active region}
which contains the elements of the array, and a rightmost region of empty
space. The array only grows into the rightmost region, and shrinks from the
leftmost region, hence one-way. The PMA is of size $m = \Theta(N)$, subject
to $m$ being a power of $2$, where $N$ is the number of elements.

When the \textit{active region} becomes too large/small due to
inserts/deletes, the PMA is rebalanced. A new array is allocated, and the
elements are spread out evenly over the new array (to minimize element
density).

Each leaf in the static CO B-tree is a pointer into a $\Theta(\log N)$-size
region in the packed-memory array.

\subsubsection{Operations}

\Search{} works as follows. Search the static tree for $k$ until a leaf node
is reached. Follow the leaf node pointer into the PMA. Scan right in the PMA
until either $k$ is located, or some element $> k$ is reached. Do not acquire
any locks during the search.

\Insert{} works as follows. Use \Search{} to locate the $\Theta(\log
N)$-region of the PMA that would contain $k$. Let $s$ be the slot
such that slot $s - 1$ contains the largest key smaller than $k$.
Lock the current region and scan right to find $s$. If scanning
forward requires going into the next $\Theta(\log N)$-region, lock the next
region and release the current region (\textit{hand-over-hand} locking
technique). Once $s$ is located, if $s$ is free, just insert
$k$. Otherwise, a \textit{rebalance} operation is required.

To perform a rebalance operation, we scan right (locking as necessary) until the
region from $s$ to the current position $s'$ is not "too dense". The
notion of density is defined in detail in \cite{BenderFiGi05}. The region $[s,
s']$ is then rebalanced as follows. Start from the rightmost element, and move
elements to the right (maintaining order) such that no suffix of $[s, s']$ is
too dense. This rebalancing creates room for $k$; we insert $k$ and then
release the locks.

\subsubsection{Complexity}

% TODO: can somebody check this time-bound? seems kind of bad

Sequential search memory transfer bounds are not presented in the paper, but presumably
since search is a static CO tree lookup followed by a scan, the time bound is
$O(\log_{B}{N} + ({\log{N}})/{B} + 1)$.

Sequential insert is $O(\log_{B}{N} + (\log^2{N})/{B} + 1)$. The proof is based
on an accounting argument that places $\Theta((\log^2{m})/B)$ dollars in each
array slot. The proof is not presented in any detail in \cite{BenderFiGi05}, but
the authors argue that it is simply a slightly-modified version of the proof
presented in \cite{Katriel02}.

\subsubsection{Lock-free Variant}

The lock-free variant is based on the PMA CO B-tree, except the locks acquired
during the PMA modification are replaced by augmenting each element in the PMA
with a marker. The marker is used to denote the ongoing (concurrent) operation of
a particular cell. The authors assume that a marker is able to fit within a
single machine-word, so that the atomic operations are available and
performant.

In addition, four non-blocking primitives are introduced:
\begin{inparaenum}[(a)]
  \item \textit{move},
  \item \textit{cell-insertion},
  \item \textit{cell-deletion}, and
  \item \textit{read}.
\end{inparaenum} These primitives are self-explanatory, except they
all work in a non-blocking manner. They can fail at any point, and it is
up to the current process to restart the operation on failure.

The concurrency protocol works as follows. Any of the non-blocking primitives
are initiated by denoting the cell in question with the marker. If a process
ever encounters a cell with a marker, it temporarily suspends its current task and
helps to complete the pending operation. When the originating process is
context-switched back in, any store-conditionals (SCs) will fail, and then the process can check to
see if the operation has already been completed. The implementation of the
non-blocking primitives is described in more detail in \cite{BenderFiGi05},
but the basic idea is using LL/SC to detect when concurrent operations are
on-going.

\begin{figure*}[h!t]
\begin{center}
    \begin{tabular}{ |   l   |  l     |     l     |     }
    \hline
	Data Structure & Search & Insert / Delete  \\ \hline \hline
	 \textit{Original CO B-tree}  & $O(\log_{B}{N})$ & $O(\log_{B}{N} + \frac{\log^2{N}}{B})$  \\ \hline
	\textit{String CO B-tree}  & $O(1+\log_{B}N+||k||/B)$ &   $O(1+\log_{B}N+\log^{2}N||k||/B)$  \\ \hline
	\textit{Streaming Shuttle Tree}  & $O(\log_{B}{N})$ &  $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$  \\ \hline
	\textit{Streaming COLA}  & $O(\log N)$ & $O((\log N) / B)$  \\ \hline
	\textit{Concurrent Exponential CO B-tree} &   $O(\log_B{N} + \log_{\alpha}{\log{B}})$  &   $O(\log_B{N} + \log{\alpha}{\log{B}})$  \\ \hline
	\textit{Concurrent PMA CO B-tree}  & $O(\log_{B}{N} + ({\log{N}})/{B} + 1)$ &   $O(\log_{B}{N} + (\log^2{N})/{B} + 1)$  \\ \hline
    \end{tabular}
    \caption{A summary of the surveyed CO B-trees. The bounds are in terms of the number of memory transfers. The bounds for \textit{Concurrent Exponential CO B-tree} are in expectation only.}
    \label{fig:table}
\end{center}
\end{figure*}

\subsubsection{Lock-free Variant Operations}

\Search{} is un-modified from the lock-based version.

\Insert{} is modified as follows. Cell $s$ is located as before. A
\textit{cell-insertion} operation is performed on $s$. If it succeeds, we are
done. Otherwise, a rebalance is necessary. The rebalance proceeds as
before, scanning right to find the region $[s, s']$. However, whenever a
non-empty cell is encountered, a load-link (LL) is performed on the cell. Once
the region is found, the rebalancing-to-the-right is performed as before.
However, a store-conditional (SC) is performed on the destination of each
cell. If any SC fails, then this means concurrent modification had occurred,
in which case the algorithm restarts from the beginning.

\section{Conclusion}

This paper presented a survey on different types of cache-oblivious B-trees: the 
original \textit{cache-oblivious B-tree}, the \textit{cache-oblivious string
B-tree}, the \textit{cache-oblivious streaming B-tree}, and the \textit{cache-oblivious
concurrent B-tree}. A summary of these data structures is provided
in Figure~\ref{fig:table}.

% are we putting this in in the final copy?
%\section{Acknowledgments}
%We sincerely thank Dr. David Karger and all of the TAs for 6.854 
f%or their guidance, help, and support. May the Flow (Max-Flow, that is) be with you all.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}

\bibliography{citations}  % sigproc.bib is the name of the Bibliography in this case

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
%%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%% This next section command marks the start of
%% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The acm\_proc\_article-sp document class file itself is chock-full of succinct
%and helpful comments.  If you consider yourself a moderately
%experienced to expert user of \LaTeX, you may find reading
%it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}